### 2024-03-25 Latex/draft.tex
add latex support

\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,amsfonts,hyperref}

\title{Roughest of Drafts}
\author{Jason Vasquez \and Dylan Skinner \and Jeff Hansen \and Benjamin McMullin}

\begin{document}

\maketitle

\begin{abstract}
    Predicting NHL games is hard. Here is our attempt
\end{abstract}

\section{Problem Statement and Motivation}
% Your content for this section goes here

\section{Data}
% Your content for this section goes here

\section{Methods}
% Your content for this section goes here

\section{Results}
% Your content for this section goes here

\section{Analysis}
% Your content for this section goes here

\section{Ethical Considerations}
% Your content for this section goes here

\section{Conclusions}
% Your content for this section goes here

% Bibliography
\bibliographystyle{plain}
\bibliography{references} % Replace 'references' with the name of your .bib file

\end{document}

### 2024-03-25 py_files/data_manager.py
Pushing data_manager file with combined states.

@@ -1,5 +1,6 @@
 import pandas as pd
 import numpy as np
+from tqdm.auto import tqdm
 
 def get_data(file):
     
@@ -23,10 +24,11 @@ def get_data(file):
 
     # Iterate through the games
     final_df = pd.DataFrame()
+    game_bar = tqdm(total=len(unique_game_ids))
     for game_id in unique_game_ids:
-        home_df, away_df = get_data(df, game_id)
-        final_df = final_df.append(home_df)
-        final_df = final_df.append(away_df)
+        final_place = get_data(df, game_id)
+        final_df = final_df.append(final_place)
+        game_bar.update(1)
 
 
 def create_state_space(df, game_id):
@@ -40,7 +42,6 @@ def create_state_space(df, game_id):
     Returns:
     tuple: A tuple containing two DataFrames, one for home team events and one for away team events.
     """
-
     # Filter the DataFrame to get data for the specified game and drop rows with NaN values in event_team_type
     game1 = df[df.game_id == game_id] 
     game1 = game1.dropna(subset=['event_team_type'])
@@ -51,12 +52,12 @@ def create_state_space(df, game_id):
     event_types = np.append(event_types, ['TIME_REMAINING', 'HOME', 'WIN', 'TEAM', 'GAME_ID', 'CORSI', 'FENWICK'])
 
     # Create dictionaries to store event counts for home and away teams
-    home_dict = {event: 0 for event in event_types}
-    away_dict = {event: 0 for event in event_types}
+    final_dict = {f'HOME_{event}': 0 for event in event_types}
+    away_dict = {f'AWAY_{event}': 0 for event in event_types}
+    final_dict.update(away_dict)
 
     # Create DataFrames to store event data for home and away teams
-    home_df = pd.DataFrame(columns=home_dict.keys())
-    away_df = pd.DataFrame(columns=away_dict.keys())
+    final_df = False
 
     # Iterate through the events in the game and count them
     for _, row in game1.iterrows():
@@ -66,39 +67,46 @@ def create_state_space(df, game_id):
         
         # Determine if the event belongs to the home or away team and update counts accordingly
         if row['event_team_type'] == 'home':
-            home_dict[row['event_type']] += 1
-            home_dict['TIME_REMAINING'] = row['game_seconds_remaining']
[…]

### 2024-03-25 py_files/data_manager.py
Fixed create_state_space (now completely combined).

@@ -30,6 +30,8 @@ def get_data(file):
         final_df = final_df.append(final_place)
         game_bar.update(1)
 
+    return final_df
+
 
 def create_state_space(df, game_id):
     """
@@ -81,12 +83,11 @@ def create_state_space(df, game_id):
                 final_dict['HOME_CORSI_FOR'] = calculate_corsi_for(final_dict['HOME_CORSI'], final_dict['AWAY_CORSI'])
                 final_dict['HOME_FENWICK_FOR'] = calculate_fenwick_for(final_dict['HOME_FENWICK'], final_dict['AWAY_FENWICK'])
             
-            # home_df = home_df.append(home_dict, ignore_index=True)
         else:
-            final_dict[f"HOME_{row['event_type']}"] += 1
+            final_dict[f"AWAY_{row['event_type']}"] += 1
             final_dict['TIME_REMAINING'] = row['game_seconds_remaining']
-            final_dict['HOME'] = 0
-            final_dict['WIN'] = 1 if row['home_final'] < row['away_final'] else 0
+            final_dict['AWAY_HOME'] = 0
+            final_dict['WIN'] = 0 if row['home_final'] < row['away_final'] else 1
             final_dict['AWAY_TEAM'] = row['team_encoded']
             final_dict['GAME_ID'] = game_id
             
@@ -97,14 +98,11 @@ def create_state_space(df, game_id):
                 final_dict['AWAY_CORSI_FOR'] = calculate_corsi_for(final_dict['AWAY_CORSI'], final_dict['HOME_CORSI'])
                 final_dict['AWAY_FENWICK_FOR'] = calculate_fenwick_for(final_dict['AWAY_FENWICK'], final_dict['HOME_FENWICK'])
                 
-            # away_df = away_df.append(away_dict, ignore_index=True)
+
         if type(final_df) is bool:
-            # print(final_dict)
             final_df = pd.DataFrame(final_dict, index=[0])
         else:
             final_df = final_df.append(final_dict, ignore_index=True)
-        
-    
         
     return final_df
 

### 2024-03-26 py_files/data_manager.py
fix state space

@@ -26,7 +26,7 @@ def get_data(file):
     final_df = pd.DataFrame()
     game_bar = tqdm(total=len(unique_game_ids))
     for game_id in unique_game_ids:
-        final_place = get_data(df, game_id)
+        final_place = create_state_space(df, game_id)
         final_df = final_df.append(final_place)
         game_bar.update(1)
 

### 2024-03-26 Latex/draft.tex
basic edits of latex paper.

@@
 
-\title{Roughest of Drafts}
+\title{Puck Predictions: Unraveling the NHL Game Forecasting Riddle}
 \author{Jason Vasquez \and Dylan Skinner \and Jeff Hansen \and Benjamin McMullin}
@@
 \begin{abstract}
-    Predicting NHL games is hard. Here is our attempt
+    The goal of this project is simple: predict the outcomes of NHL games from any given state.
+    As simple as the problem statement is, however, the solution is not so straightforward.
+    To solve this problem, we will use a variety of machine learning techniques, including logistic regression,
+    XGBoost, and ARIMA models. Additionally, we utilize a form of MCMC to simulate the outcomes of games from any given
+    state. Our hypothesis is that we will be able to successfully predict the outcomes of NHL games with a high degree of accuracy
+    using these tools.
 \end{abstract}
@@
 % Your content for this section goes here
+Our data came from the hockeyR Github repository\cite{hockeyR-data}. This repository contains an abundance of data about every NHL game
+that has occured since the 2010-11 season. This data includes information about the events that transpire in a game (hits, shots, goals, etc.),
+which teams are playing, who is on the ice, and the final score of the game.
 

### 2024-03-27 Latex/draft.tex

@@
 that has occured since the 2010-11 season. This data includes information about the events that transpire in a game (hits, shots, goals, etc.),
-which teams are playing, who is on the ice, and the final score of the game.
+which teams are playing, who is on the ice, and the final score of the game. The data is stored in a series of {\tt .csv.gz} files, allowing for
+easy access and manipulation.
+
+Each game in a season is given a unique identifier ({\tt game\_id}), which is constant across all events in a game. Every event that occurs in a game
+will be stored in the {\tt event\_type} column. There are 17 unique event types, including things such as game start, faceoff, shot, hit, and goal.
+Most of these event types are not relevant to our analysis, so we remove them from the dataset. After removing the unnecessary events, we are left with
+nine events: blocked shot, faceoff, giveaway, goal, hit, missed shot, penalty, shot, and takeaway. These events are attributed to the
+team and player that performs the event. We only take into consideration the team that performs the event and discard the player information.
+
+The data also contains information about when the event occured. This appears in a variaty of formats, but we only
+use the {\tt game\_time\_remaining} column. {\tt game\_time\_remaining} starts
+at 3600 (60 minutes) and counts down to 0. If the game goes into extra time, i.e., it is tied after 60 minutes, {\tt game\_time\_remaining} will
+be a negative value.
+
+We found that our data did not contain any missing values that was not easily explainable. For example, if a game is starting, there will be no
+events for penalties, which will result in a {\tt NaN} value in the penalties column. Additionally, any data that was confusing or not easily explainable
+(for example the home team having 7 players on the ice and the away team having 5), was manually verified by watching a clip of the game where
+the event occured to make sure the event was recorded correctly. We did not find any incorrectly recorded events, so we 
+did not remove any strange events from out dataset.
 
 \section{Methods}
+\subsection{t-SNE, UMAP, and PCA}
+
+\subsection{Regression and XGBoost}
+
+\subsection{MCMC Game Simulation}
+
+\subsection{Exponential Smoothing}
 % Your content for this section goes here

### 2024-03-27 Latex/draft.tex
ethics section

@@
 \documentclass[11pt]{article}
-\usepackage{amsmath,amssymb,amsthm,amsfonts,hyperref}
+\usepackage{amsmath,amssymb,amsthm,amsfonts,hyperref, enumitem}
 
@@
 \section{Methods}
+
+\subsection{Bayesian Network}
+
 \subsection{t-SNE, UMAP, and PCA}
@@
 \subsection{MCMC Game Simulation}
-
-\subsection{Exponential Smoothing}
-% Your content for this section goes here
 
@@
 \section{Ethical Considerations}
-% Your content for this section goes here
+Predicting win percentages or outcomes in hockey games, like any sport, raises several ethical considerations. Here are some key points we want to address:
+
+\begin{itemize}[label=\textbullet]
+\item \textbf{Gambling and Addiction}:
+Our win percentages and predictions might be used by those who wish to gamble, which could lead to addiction and financial harm, especially if undue trust in placed in these methods.
+Any publication of these methods or predictions would be accompanied by promoting healthy and responsible gambling practices.
+
+\item \textbf{Fairness and Integrity of the Game}:
+Sometimes, coaches and players becoming awareness of their 'chance' of winning can affect how the game is played, potentially harming the integrity of the sport.
+We must be careful to not provide an unfair advantage to any team or player.
+Inaccurate predictions could lead a team to believe that the game is out of reach when it isn't, and we want to avoid that.
+
+\end{itemize}
+
+Overall, our predictions, like any, should not be considered declarative for gambling or performance purposes, but are rather an interesting exposition into the complexity of sporting events.
 

### 2024-03-27 Latex/draft.tex

@@
 \subsection{Bayesian Network}
+We first used a Bayesian Network to establish a benchmark for probability using several key features.
+
+Bayesian networks are probabilistic graphical models that represent probabilistic relationships among a set of variables using a directed acyclic graph (DAG). 
+In a Bayesian network, nodes represent random variables, and directed edges between nodes represent probabilistic dependencies between the variables.
+Each node in the graph is associated with a conditional probability distribution that quantifies the probability of that variable given its parent variables in the graph.
+
+For our purposes, we predefined the structure of the network, and used the data to calculate the conditional probabilities for each node.
+We then used the network to calculate the probability of a team winning given the current state of the game.
+
+The computational complexity of Bayesian Network inference is high, with exact inference being an NP-hard problem TODO CITE. 
+Using the python package pgmpy, we originally tried to fit a network with all 26 of our features, but our computational resources failed to fit this network.
+Then, to get a baseline for our future predictions, we simply fitted the model with the base features of home goals, away goals, home shots, away shots, home blocked shots, and away blocked shots. 
+These features were chosen as priors because of our opinion that they are the most important to the game, based upon our knowledge of hockey.
 
@@
 \item \textbf{Fairness and Integrity of the Game}:
-Sometimes, coaches and players becoming awareness of their 'chance' of winning can affect how the game is played, potentially harming the integrity of the sport.
+Sometimes, coaches and players becoming aware of their chance of winning can affect how the game is played, potentially harming the integrity of the sport.
 We must be careful to not provide an unfair advantage to any team or player.

### 2024-03-28 Latex/draft.tex

@@
 % Your content for this section goes here
+In the world of sports analytics, predicting the outcomes of games is a common and challenging problem, with live win predictions adding
+an extra layer of complexity. For most sports, there are a plethora of widely accepted—yet hidden—predictive models and methods that are used to
+predict games. In addition to this, most sports have easily accessible statistics and graphics that give current win probabilities for any live game.
+
+Hockey, however, is a different story. While there are some methods used to predict the outcome of National Hockey League (NHL) games, these models
+typically belong to sport books and their nuances are not publicly disclosed. Additionally, hockey analytics is not as
+developed as it is in other sports, such as basketball or baseball. This lack of model transparency and public interest in hockey analytics
+makes predicting the outcomes of NHL games a challenging problem. That is where this project comes in. The goal of this project is simple: predict
+the outcomes of NHL games and provide a live win probability for any given game state.
+
+Despite the simplicity of the problem statement, as mentioned, the solution is not so straightforward. The NHL provides fast-paced games with many events
+occuring in quick succession. Our goal is to sift through the data points and identify key features that will allow us to predict the outcome of a game and
+ignore other features that exist as seeming proxy noise.
+
+Our motivation for this project exists strictly as fans of the sport and as data scientists. Our model is not intended to be used for gambling or any other
+nefarious purposes—any use of this model for such purposes is a misuse of our work.
 

### 2024-03-28 Latex/draft.tex
Added basic DAG to see how it works.
Worked on problem statement and motivation.

@@
 \documentclass[11pt]{article}
-\usepackage{amsmath,amssymb,amsthm,amsfonts,hyperref, enumitem}
+\usepackage{amsmath,amssymb,amsthm,amsfonts,hyperref, enumitem, tikz}
 
@@
 developed as it is in other sports, such as basketball or baseball. This lack of model transparency and public interest in hockey analytics
-makes predicting the outcomes of NHL games a challenging problem. That is where this project comes in. The goal of this project is simple: predict
-the outcomes of NHL games and provide a live win probability for any given game state.
+makes predicting the outcomes of NHL games a very underdeveloped and challenging problem. Previous attempts and research into predicting NHL games
+has relied on methods such as decision trees and artificial neural networks~\cite{pishcedda} (from 2014), naïve bayes and support vector machines~\cite{weissbock2013use} (from 2013),
+and Monte Carlo simulations~\cite{Weissbock2014ForecastingSI} (from 2014).
 
-Despite the simplicity of the problem statement, as mentioned, the solution is not so straightforward. The NHL provides fast-paced games with many events
-occuring in quick succession. Our goal is to sift through the data points and identify key features that will allow us to predict the outcome of a game and
-ignore other features that exist as seeming proxy noise.
+In addition to model research, some research has also gone into developing new features that can be used to better predict the game outcomes. 
+The two biggest engineered classes of features are the Corsi (\url{https://en.wikipedia.org/wiki/Corsi_(statistic)})
+ and Fenwick (\url{https://en.wikipedia.org/wiki/Fenwick_(statistic)}) metrics (both from 2007).
+
+Our project seeks a similar outcome to the research mentioned above: predict the outcomes of NHL games. Not only this,
+but we seek to provide live, acturate win probabilities for any given game state. Despite the simplicity of the problem statement, 
+as mentioned, the solution is not so straightforward. The NHL provides fast-paced games with many events
+occuring in quick succession. Our goal is to use this abundance of data and new approaches to build upon previous research.
 
@@
 
+(Basic graph to test)
+
+\begin{center}
+    \begin{tikzpicture}[node distance={15mm}, thick, main/.style = {draw, circle}]
+        \node[main] (1) {t r}; 
+        \node[main] (2) [below left of=1] {t g};
+        \node[main] (3) [below right of=1] {t s};
+        \node[main] (4) [below right of=3] {t b};
+
+        \draw[->] (1) -- (2);
+        \draw[->] (1) -- (3);
+        \draw[->] (1) to [out=15,in=90,looseness=1.5] (4);
+
+    \end{tikzpicture} 
+\end{center}
+
 \subsection{t-SNE, UMAP, and PCA}

### 2024-03-28 py_files/data_manager.py
added corrected strength_state columns

@@ -85,7 +85,22 @@ def load_and_clean_csv(path):
     else:
         df['date_time'] = "-"
         df['game_date'] = "-"
-            
+        
+    # setup the strenght column to always be in the perspective of home
+    df['home_skaters'] = df['home_skaters'].astype(int)
+    df['away_skaters'] = df['away_skaters'].astype(int)
+    df['home_skaters'] = df['home_skaters'].replace(0, np.nan).ffill().astype(int)
+    df['away_skaters'] = df['away_skaters'].replace(0, np.nan).ffill().astype(int)
+    df['strength_state'] = df['home_skaters'].copy().astype(str) + 'v' + df['away_skaters'].copy().astype(str)
+    
+    df['strength_code'] = 'EV'
+    df.loc[df['home_skaters'] > df['away_skaters'], 'strength_code'] = 'PP'
+    df.loc[df['home_skaters'] < df['away_skaters'], 'strength_code'] = 'SH'
+    
+    df['strength'] = 'Even'
+    df.loc[df['home_skaters'] > df['away_skaters'], 'strength'] = 'Power Play'
+    df.loc[df['home_skaters'] < df['away_skaters'], 'strength'] = 'Short Handed'
+    
     return df
 
 def load_clean_feather(year):

### 2024-03-28 Latex/draft.tex
syncing
Added footnote to draft for corsi, fenwick.

@@
 In addition to model research, some research has also gone into developing new features that can be used to better predict the game outcomes. 
-The two biggest engineered classes of features are the Corsi (\url{https://en.wikipedia.org/wiki/Corsi_(statistic)})
- and Fenwick (\url{https://en.wikipedia.org/wiki/Fenwick_(statistic)}) metrics (both from 2007).
+The two biggest engineered classes of features are the Corsi
+ and Fenwick\footnote{These metrics were created by sports bloggers Tim Barnes and Mark Fenwick, respectively. 
+ We were unable to locate the original blog posts talking about these metrics, but a good article to learn more about the math can be
+ found here \url{https://thehockeywriters.com/corsi-fenwick-stats-what-are-they/}.} metrics (both around 2007).
 

### 2024-03-28 py_files/simulator.py
syncing

@@ -1,0 +1,1 @@
+def simulate_games(curr_state_df)

### 2024-03-28 Latex/draft.tex
worked on MCMC latex

@@
 \usepackage{amsmath,amssymb,amsthm,amsfonts,hyperref, enumitem, tikz}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 
+To simulate hockey games, we setup a Markov Chain where the states are a tuple of three consecutive
+events that occurred. For example, if the home team won a faceoff, the home team lost the puck, and 
+then the away team shot the puck and missed, the markov chain would look like figure \ref{fig:markov_chain_sample}.
+The probabilities of transitioning from one triple-state to another triple-state is calculated by:
+\begin{align*}
+    P(s_{t+1} = (B,C,D) \;|\; s_t = (A,B,C)) &= P(D \;|\; (A,B,C)) \\
+    &= \frac{\{ \text{\# of times (A,B,C,D) happend}\}}{\{ \text{\# of times (A,B,C) happened}\}} \\
+\end{align*}
+Where $A,B,C,D$ are events that can occur in a game, and the tuple (A,B,C,D) represents that "A then 
+B then C then D" happend right after eachother in a game.
+
+To simulate a game, we performed this monte carlo algorithm:
+
+\begin{algorithm}
+\caption{Simulation Algorithm}
+\begin{algorithmic}[1]
+\State $\text{time\_remaining} \gets 3600$
+\State $s0 \gets "\#"$
+\State $s1 \gets "\#"$
+\State $s2 \gets "\#"$
+
+\While{$\text{time\_remaining} > 0$}
+    \State $\text{next\_state} \gets \text{sample from } \text{MC(curr\_state)}$
+    \State $s0 \gets s1$
+    \State $s1 \gets s2$
+    \State $s2 \gets \text{next\_state}$
+
+    \State $\text{event\_time} \gets \text{sample from } \text{KDE\_times()}$
+    \State $\text{time\_remainting} \gets \text{time\_remainting} - \text{event\_time}$
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
+
+% add the image images/markov_chain_sample.png
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.8\textwidth]{images/markov_chain_sample.png}
+    \caption{Sample Markov Chain for the current state (home team won faceoff, 
+    home team lost puck, away team missed shot) with the probabilities of the next event 
+    sorted from highest to lowest}
+    \label{fig:markov_chain_sample}
+\end{figure}
+
 \section{Results}

### 2024-03-28 Latex/draft.tex

@@
 
-To simulate hockey games, we setup a Markov Chain where the states are a tuple of three consecutive
+To simulate hockey games, we created a Markov Chain where the states are a tuple of three consecutive
 events that occurred. For example, if the home team won a faceoff, the home team lost the puck, and 
 then the away team shot the puck and missed, the markov chain would look like figure \ref{fig:markov_chain_sample}.
-The probabilities of transitioning from one triple-state to another triple-state is calculated by:
-\begin{align*}
-    P(s_{t+1} = (B,C,D) \;|\; s_t = (A,B,C)) &= P(D \;|\; (A,B,C)) \\
-    &= \frac{\{ \text{\# of times (A,B,C,D) happend}\}}{\{ \text{\# of times (A,B,C) happened}\}} \\
-\end{align*}
-Where $A,B,C,D$ are events that can occur in a game, and the tuple (A,B,C,D) represents that "A then 
-B then C then D" happend right after eachother in a game.
 
-To simulate a game, we performed this monte carlo algorithm:
-
-\begin{algorithm}
-\caption{Simulation Algorithm}
-\begin{algorithmic}[1]
-\State $\text{time\_remaining} \gets 3600$
-\State $s0 \gets "\#"$
-\State $s1 \gets "\#"$
-\State $s2 \gets "\#"$
-
-\While{$\text{time\_remaining} > 0$}
-    \State $\text{next\_state} \gets \text{sample from } \text{MC(curr\_state)}$
-    \State $s0 \gets s1$
-    \State $s1 \gets s2$
-    \State $s2 \gets \text{next\_state}$
-
-    \State $\text{event\_time} \gets \text{sample from } \text{KDE\_times()}$
-    \State $\text{time\_remainting} \gets \text{time\_remainting} - \text{event\_time}$
-\EndWhile
-\end{algorithmic}
-\end{algorithm}
-
-% add the image images/markov_chain_sample.png
-\begin{figure}[ht]
+\begin{figure}[H]
     \centering
@@
     \label{fig:markov_chain_sample}
+\end{figure}
+
+The probabilities of transitioning from one triple-state to another triple-state is calculated by:
+\begin{align*}
+    P(s_{t+1} = (B,C,D) \;|\; s_t = (A,B,C)) &= P(D \;|\; (A,B,C)) \\
+    &= \frac{\{ \text{\# of times (A,B,C,D) happend}\}}{\{ \text{\# of times (A,B,C) happened}\}} \\
+\end{align*}
+Where $A,B,C,D$ represent events that can occur in a game, and the tuple (A,B,C,D) represents that "A then 
+B then C then D" happened right after each other in a game.
+
+To simulate a game, we performed this Monte Carlo algorithm that acted like a random walk through the Markov Chain:
+\begin{algorithm}[H]
+    \caption{Simulation Algorithm}
+    \begin{algorithmic}[1]
+        \State $\text{time\_remaining} \gets 3600$
+        \Comment{NHL games are 3600 seconds}
+        \State $s_0 \gets "\#"$
+        \Comment{The "\#" symbol represents the start of a game}
+        \State $s_1 \gets "\#"$
+        \State $s_2 \gets "\#"$
+        %\State $\text{event\_counts} \gets \text{dictionary()}$
+
+        \While{$\text{time\_remaining} > 0$}
+            \State $\text{next\_state} \gets \text{sample from } \text{MC(curr\_state)}$
+            %\State $\text{event\_counts}[\text{next\_state}] \mathrel{+}= 1$
+
+            \State $s_0 \gets s_1$
+            \State $s_1 \gets s_2$
+            \State $s_2 \gets \text{next\_state}$
+
+            \State $\text{event\_time} \gets \text{sample from } \text{KDE\_times()}$
+            \State $\text{time\_remaining} \gets (\text{time\_remaining} - \text{event\_time})$
+        \EndWhile
+    \end{algorithmic}
+    \label{alg:monte_carlo_algorithm}
+\end{algorithm}
+The KDE\_times() is a KDE model fit on the amount of seconds that transpired between each hockey event for all NHL games over the course of 13 years.
+This KDE model resembled a poisson distribution with $\lambda \approx 18$.
+During the course of the algorithm, we maintained a count of all of the events that transpired.
+Upon termination of the simulation, we compared home goals to away goals and termed a winner.
+If there was a tie, we would run the algorithm until another team scored a goal, thus simulating overtime per NHL rules.
+
+To predict a probability of a certain team winning a hockey game, we would run 50 simulations with our algorithm, but we would initialize the starting states $(s_0, s_1, s_2)$ to be the most current events in the hockey game.
+Out of the 50 simulations, we would compute $P(\text{home winning}) = \{\text{\# home simulation wins}\}/50$ and $P(\text{away winning}) = 1-P(\text{home winning})$.
+
+To evaluate our simulation's effectiveness at accurately modeling actual hockey games, we combined a dataset with the final event counts for 1500 actual hockey games and for 1500 simulated games.
+We then performed PCA, t-SNE and UMAP dimensionality reductions using various perplexity and neighbor hyperparameters to see if these algorithms clustered the synthetic games and actual games in different clusters.
+As shown below in \ref{fig:simulation_v_actual}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{images/simulation_v_actual_dimred.png}
+    \caption{PCA, t-SNE and UMAP performed on a dataset with final event counts for 1500 simulated and 1500 actual NHL games. This evidences that our simulation accurately emulates live NHL games}
+    \label{fig:simulation_v_actual}
 \end{figure}

### 2024-03-29 Latex/draft.tex
added a few more things to MCMC LaTeX

@@
 \end{algorithm}
-The KDE\_times() is a KDE model fit on the amount of seconds that transpired between each hockey event for all NHL games over the course of 13 years.
-This KDE model resembled a poisson distribution with $\lambda \approx 18$.
-During the course of the algorithm, we maintained a count of all of the events that transpired.
-Upon termination of the simulation, we compared home goals to away goals and termed a winner.
-If there was a tie, we would run the algorithm until another team scored a goal, thus simulating overtime per NHL rules.
+The KDE\_times() is a KDE model fit on the amount of seconds that transpired between 
+each hockey event for all NHL games over the course of 13 years. This KDE model 
+resembled a poisson distribution with $\lambda \approx 18$. During the course of 
+the algorithm, we maintained a count of all of the events that transpired. Upon 
+termination of the simulation, we compared home goals to away goals and termed 
+a winner. If there was a tie, we would run the algorithm until another team scored0 
+a goal, thus simulating overtime per NHL rules.
 
-To predict a probability of a certain team winning a hockey game, we would run 50 simulations with our algorithm, but we would initialize the starting states $(s_0, s_1, s_2)$ to be the most current events in the hockey game.
-Out of the 50 simulations, we would compute $P(\text{home winning}) = \{\text{\# home simulation wins}\}/50$ and $P(\text{away winning}) = 1-P(\text{home winning})$.
+To predict a probability of a certain team winning a hockey game, we would run 
+50 simulations with our algorithm, but we would initialize the starting states 
+$(s_0, s_1, s_2)$ to be the most current events in the hockey game. Out of the 
+50 simulations, we would compute 
+$P(\text{home winning}) = \{\text{\# home simulation wins}\}/50$ and 
+$P(\text{away winning}) = 1-P(\text{home winning})$.
 
-To evaluate our simulation's effectiveness at accurately modeling actual hockey games, we combined a dataset with the final event counts for 1500 actual hockey games and for 1500 simulated games.
-We then performed PCA, t-SNE and UMAP dimensionality reductions using various perplexity and neighbor hyperparameters to see if these algorithms clustered the synthetic games and actual games in different clusters.
-As shown below in \ref{fig:simulation_v_actual}
+To evaluate our simulation's effectiveness at accurately modeling actual hockey 
+games, we combined a dataset with the final event counts for 1500 actual hockey 
+games and for 1500 simulated games. We then performed PCA, t-SNE and UMAP 
+dimensionality reductions using various perplexity and neighbor hyperparameters 
+to see if these algorithms clustered the synthetic games and actual games in 
+different clusters. As shown below in \ref{fig:simulation_v_actual}, the simulated
+games and actual games are all clustered together, thus demonstrating that our
+simulation effectively emulates live NHL games.
 

### 2024-03-29 Latex/draft.tex
add dag picture

@@
 Using the python package pgmpy, we originally tried to fit a network with all 26 of our features, but our computational resources failed to fit this network.
-Then, to get a baseline for our future predictions, we simply fitted the model with the base features of home goals, away goals, home shots, away shots, home blocked shots, and away blocked shots. 
+Then, to get a baseline for our future predictions, we simply fitted the model with the base features of time remaining (tr), home goals (hg), away goals (ag), home shots (hs), away shots (as), home blocked shots (hbs), and away blocked shots (abs).
 These features were chosen as priors because of our opinion that they are the most important to the game, based upon our knowledge of hockey.
 
-(Basic graph to test)
+The conditional dependencies of the chosen network are shown in the DAG below:
 
@@
     \begin{tikzpicture}[node distance={15mm}, thick, main/.style = {draw, circle}]
-        \node[main] (1) {t r}; 
-        \node[main] (2) [below left of=1] {t g};
-        \node[main] (3) [below right of=1] {t s};
-        \node[main] (4) [below right of=3] {t b};
+        \node[main] (1) {tr}; 
+        \node[main] (2) [below left of=1] {hs};
+        \node[main] (3) [below right of=1] {as};
+        \node[main] (4) [below right of=3] {hbs};
+        \node[main] (5) [below left of=2] {abs};
+        \node[main] (6) [below right of=5] {hg};
+        \node[main] (7) [below left of=4] {ag};
 
@@
         \draw[->] (1) -- (3);
-        \draw[->] (1) to [out=15,in=90,looseness=1.5] (4);
+        \draw[->] (3) -- (4);
+        \draw[->] (2) -- (5);
+        \draw[->] (5) -- (6);
+        \draw[->] (4) -- (7);
+        \draw[->] (3) -- (7);
+        \draw[->] (2) -- (6);
 
@@
     \includegraphics[width=0.8\textwidth]{images/simulation_v_actual_dimred.png}
-    \caption{PCA, t-SNE and UMAP performed on a dataset with final event counts for 1500 simulated and 1500 actual NHL games. This evidences that our simulation accurately emulates live NHL games}
+    \caption{PCA, t-SNE and UMAP performed on a dataset with final event counts for 1500 simulated and 1500 actual NHL games.
+The joined grouping demonstrates that our simulation accurately emulates live NHL games}
     \label{fig:simulation_v_actual}

### 2024-03-29 Latex/draft.tex
Added bayesian network citation and edited tsne graph.

@@
 % Your content for this section goes here
-Our data came from the hockeyR Github repository\cite{hockeyR-data}. This repository contains an abundance of data about every NHL game
+Our data came from the hockeyR Github repository~\cite{hockeyR-data}. This repository contains an abundance of data about every NHL game
 that has occured since the 2010-11 season. This data includes information about the events that transpire in a game (hits, shots, goals, etc.),
@@
 
-The computational complexity of Bayesian Network inference is high, with exact inference being an NP-hard problem TODO CITE. 
+The computational complexity of Bayesian Network inference is high, with exact inference being an NP-hard problem~\cite{pmlr-vR0-chickering95a}. 
 Using the python package pgmpy, we originally tried to fit a network with all 26 of our features, but our computational resources failed to fit this network.
@@
 \end{center}
-
-\subsection{t-SNE, UMAP, and PCA}
 
@@
 to see if these algorithms clustered the synthetic games and actual games in 
-different clusters. As shown below in \ref{fig:simulation_v_actual}, the simulated
+different clusters. As shown below in~\ref{fig:simulation_v_actual}, the simulated
 games and actual games are all clustered together, thus demonstrating that our
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{images/simulation_v_actual_dimred.png}
-    \caption{PCA, t-SNE and UMAP performed on a dataset with final event counts for 1500 simulated and 1500 actual NHL games.
+    \includegraphics[width=0.9\textwidth]{images/pca_tsne_umap_sim_act.png}
+    \caption{PCA, t-SNE, and UMAP performed on a dataset with final event counts for 1500 simulated and 1289 actual NHL games.
 The joined grouping demonstrates that our simulation accurately emulates live NHL games}

### 2024-03-29 simulate_games.py
syncing

@@ -121,6 +121,85 @@ def simulate_games(n_games:int, save_path:str):
 
     return games_full
 
+def get_simulation_prob(curr_snapshot:pd.DataFrame, probs:pd.DataFrame, kde):
+    
+    # extract the events_dictionary, seconds_remaining, and the latest
+    # three events from the current game snapshot
+    curr_dict = curr_snapshot.to_dict(orient='records')[-1]
+    seconds_remaining = curr_dict['game_seconds_remaining']
+    
+    diffs = (curr_snapshot.diff()
+        .dropna()
+        .reset_index(drop=True)
+        .drop(columns=['game_seconds_remaining']))
+    cols = np.array(diffs.columns)
+    inds = np.argmax(diffs.to_numpy(), axis=1)
+    events = cols[inds]
+    if len(events) >= 3:
+        prev3, prev2, prev1 = events[-3], events[-2], events[-1]
+    elif len(events) == 2:
+        prev3, prev2, prev1 = '#', events[-2], events[-1]
+    elif len(events) == 1:
+        prev3, prev2, prev1 = '#', '#', events[-1]
+    else:
+        prev3, prev2, prev1 = '#', '#', '#'
+    
+    ####################################
+    #        SIMULATE THE GAME         #
+    ####################################
+    
+    n_games = 50
+
+    # simulte n_games
+    games = []
+    last_game_states = []
+    game_bar = tqdm(total=n_games)
+    for game_id in range(n_games):
+        
+        i = 0
+        samples = kde.sample(100000).astype(int).flatten()
+        
+        # setup the start of the game
+        game_id = str(game_id).zfill(8)
+        
+        # the curr_dict will keep track of the current state of the game
+        # with counts of each event
+        curr_dict['time_remaining'] = seconds_remaining
+        game_dicts = [curr_dict.copy()]
+        
+        while ((seconds_remaining > 0) or 
+               (curr_dict['GOAL_HOME'] == curr_dict['GOAL_AWAY'])):
+            
+            game_bar.set_description(str(seconds_remaining))
+            
+            # select a next event based on the probabilities
+            curr_table = probs[(probs['prev3'] == prev3) & (probs['prev2'] == prev2) & (probs['prev1'] == prev1)]
+            curr_event = np.random.choice(curr_table['curr_event'], p=curr_table['probability_avg'])
+            prev3, prev2, prev1 = prev2, prev1, curr_event
+            
+            # sample from the distribution of how long it takes for an
[…]

### 2024-03-29 Latex/draft.tex
syncing
finish bayesian method

@@
 Using the python package pgmpy, we originally tried to fit a network with all 26 of our features, but our computational resources failed to fit this network.
-Then, to get a baseline for our future predictions, we simply fitted the model with the base features of time remaining (tr), home goals (hg), away goals (ag), home shots (hs), away shots (as), home blocked shots (hbs), and away blocked shots (abs).
+Then, to get a baseline for our future predictions, we simply fitted the model with the base features of time remaining (tr), home goals (hg), away goals (ag), home shots (hs), away shots (as), home blocked shots (hbs), and away blocked shots (abs) in order to predict wins (w).
 These features were chosen as priors because of our opinion that they are the most important to the game, based upon our knowledge of hockey.
@@
         \node[main] (7) [below left of=4] {ag};
+        \node[main] (8) [below left of=7] {w};
 
@@
         \draw[->] (2) -- (6); 
+        \draw[->] (6) -- (8);
+        \draw[->] (7) -- (8);
 
@@
 \end{center}
+
+This model was chosen for the task because different stats in hockey are conditionally dependent of each other, so by modeling those conditional
+dependencies and feeding them into the model, we can hopefully acheive a more accurate prediction of the outcome of the game.
 

### 2024-03-29 Latex/draft.tex
some closer stuff

@@
 
-\section{Results}
+\section{Results and Analysis}
 % Your content for this section goes here
 
-\section{Analysis}
-% Your content for this section goes here
+\subsection{Bayesian Network}
+
+Overall, the Bayesian Network performed well, and was able to produce realistic win probabilities. Because it defined a joint distribution using a DAG and
+then used that distribution to fit the data, it was able to correctly predict accuracies using the few features provided. However, the Bayesian Network struggled to capture the intricate dependencies
+of factors other than goals that could affect the win probabilities, so the predicted probabilities are little more than an over probability calculation of goals and time reamining given historical data.
+We see this in the probability graph below, where the changes in probability correspond to the goals scored in the game.
 

### 2024-03-29 Latex/draft.tex
Pushing latex changes.

@@
 \usepackage{algpseudocode}
+\usepackage{subfigure}
 
@@
 
-\subsection{Regression and XGBoost}
+\subsection{XGBoost}
+
+In our NHL analysis research, we partitioned the dataset into segments 
+corresponding to individual teams' games over multiple seasons. 
+This enabled the creation of time series data in the form of a state vector, capturing the 
+play-by-play dynamics of each team's matches. We then trained 
+separate XGBoost models for each team to learn their unique patterns 
+and strategies. Leveraging the {\tt .predict\_proba()} method of XGBoost, we generated 
+probabilistic predictions at various stages of a game, allowing us to 
+plot the evolving probability of each team winning throughout the match
+(see~\ref{fig:xgboost_res}). 
+This approach provided insights into momentum shifts (when one team 
+has an advantage due to them having more players on the ice due to 
+a penalty by the opposing team), key moments (such as goals and 
+penalties, and critical plays influencing game outcomes).
+
+% Images
+\begin{figure}[H]
+    \centering
+    \subfigure[]{\includegraphics[width=0.9\textwidth]{images/good_xgb_example.png}} 
+    \subfigure[]{\includegraphics[width=0.9\textwidth]{images/good_xgb_example(2).png}} 
+    \caption{For both of these plots, the home team is above the the 50\% line, with the
+    away team below. The title has the final score of the game. \\
+        \textbf{(a)} The live win probabilities for the Nashville Predators at the Vegas Golden Knights. 
+    Notice the probability shifts after one team scores a goal.  \\
+        \textbf{(b)} The live win probabilities for the Calgary Flames at the Winnipeg Jets.
+        Notice the stark changes in probabilities after each goal. Also notice the
+        shift in probabilities even when no goals are scored. This suggests
+        the XGBoost model is able to pick up on momentum shifts.}
+    \label{fig:xgboost_res}
+\end{figure}
 
@@
 events that occurred. For example, if the home team won a faceoff, the home team lost the puck, and 
-then the away team shot the puck and missed, the markov chain would look like figure \ref{fig:markov_chain_sample}.
+then the away team shot the puck and missed, the markov chain would look like figure~\ref{fig:markov_chain_sample}.
 

### 2024-03-29 Latex/draft.tex
add graph for bayes

@@
     \centering
-    \subfigure[]{\includegraphics[width=0.9\textwidth]{images/good_xgb_example.png}} 
-    \subfigure[]{\includegraphics[width=0.9\textwidth]{images/good_xgb_example(2).png}} 
+    \subfigure[]{\includegraphics[width=0.6\textwidth]{images/good_xgb_example.png}} 
+    \subfigure[]{\includegraphics[width=0.6\textwidth]{images/good_xgb_example(2).png}} 
     \caption{For both of these plots, the home team is above the the 50\% line, with the
@@
 of factors other than goals that could affect the win probabilities, so the predicted probabilities are little more than an over probability calculation of goals and time reamining given historical data.
-We see this in the probability graph below, where the changes in probability correspond to the goals scored in the game.
+We see this in the probability graph below, where the changes in probability correspond to the goals scored in the game. A sample win probability graph is seen below in ~\ref{fig:bayesian_graph}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.9\textwidth]{images/good_bayesian_example.png}
+    \caption{Win probability graph using the Bayesian Network. As seen above, the win probability stayed rather stagnant until a goal was scored, where the probability shifted.}
+    \label{fig:bayesian_graph}
+\end{figure}
 

### 2024-03-29 py_files/simulator.py
syncing

@@ -1,1 +1,223 @@
-def simulate_games(curr_state_df)
+import pandas as pd
+import pickle
+from tqdm import tqdm
+import numpy as np
+import os
+import sys
+
+def simulate_games(n_games:int, save_path:str):
+    
+    # if the save_path is not a valid save path, raise a ValueError
+    dir_name = os.path.dirname(save_path)
+    if not os.path.exists(dir_name):
+        raise ValueError(
+            f"{dir_name} does not exist, so save_path={save_path} not valid")
+
+    # load the average probability table
+    probs = pd.read_csv("data/probabilities_avg_NOHIT.csv")
+
+    # more or less normalize the probabilities
+    probs_sums = (probs
+        .copy()
+        .groupby(by=['prev3', 'prev2', 'prev1'])
+        [['probability_avg', 'probability_1', 'probability_2', 'probability_3']]
+        .sum()
+        .reset_index()
+        .rename(columns={
+            'probability_avg': 'probability_avg_sum',
+            'probability_1': 'probability_1_sum',
+            'probability_2': 'probability_2_sum',
+            'probability_3': 'probability_3_sum'}))
+
+    probs2 = pd.merge(probs, probs_sums, on=['prev3', 'prev2', 'prev1'], how='left')
+    probs2['probability_avg'] /= probs2['probability_avg_sum']
+    probs2['probability_1'] /= probs2['probability_1_sum']
+    probs2['probability_2'] /= probs2['probability_2_sum']
+    probs2['probability_3'] /= probs2['probability_3_sum']
+    probs2 = probs2.drop(columns=[
+        'probability_avg_sum', 'probability_1_sum', 'probability_2_sum', 'probability_3_sum'])
+
+    probs2 = probs2.fillna(0)
+    
+    # get a list of the possible events
+    events = probs2['curr_event'].unique()
+
+    # sample 100000 times from the saved NOHIT kde
+    kde_nohit_path = "data/pickles/kde_seconds_NOHIT.pickle"
+    with open(kde_nohit_path, 'rb') as file:
+        kde = pickle.load(file)
+    
+    ####################################
+    #        SIMULATE THE GAME         #
+    ####################################
+    
+    games = []
+
+    # simulte n_games
+    game_bar = tqdm(total=n_games)
+    for game_id in range(n_games):
+        
+        i = 0
+        samples = kde.sample(100000).astype(int).flatten()
+        
+        # setup the start of the game
[…]

### 2024-03-29 simulate_games.py

@@ -4,6 +4,10 @@ from tqdm import tqdm
 import numpy as np
 import os
 import sys
+import time
+from jeffutils.utils import stack_trace
+
+from py_files.simulator import get_simulation_prob
 
 def simulate_games(n_games:int, save_path:str):
     
@@ -121,97 +125,110 @@ def simulate_games(n_games:int, save_path:str):
 
     return games_full
 
-def get_simulation_prob(curr_snapshot:pd.DataFrame, probs:pd.DataFrame, kde):
+def get_game_probabilities(game_id):
     
-    # extract the events_dictionary, seconds_remaining, and the latest
-    # three events from the current game snapshot
-    curr_dict = curr_snapshot.to_dict(orient='records')[-1]
-    seconds_remaining = curr_dict['game_seconds_remaining']
+    df_full = pd.read_feather("data/play_by_play/play_by_play_full_state_space.feather")
     
-    diffs = (curr_snapshot.diff()
-        .dropna()
+    # load the probability table
+    prob_mc = pd.read_csv("data/probability_tables/probabilities_avg_NOHIT.csv")
+    prob_mc = prob_mc[['prev3', 'prev2', 'prev1', 'curr_event', 'probability_avg']]
+    prob_mc = prob_mc.rename(columns={'probability_avg': 'probability'})
+
+    df = df_full.copy()
+
+    curr_game = df.loc[df['game_id'] == game_id, :].copy()
+    home_name = curr_game['home_name'].values[0]
+    away_name = curr_game['away_name'].values[0]
+    game_date = curr_game['game_date'].values[0]
+
+    simulation_cols = ['game_seconds_remaining'] + [c for c in curr_game.columns if 'STATE' in c]
+    simulation_cols_new = [c.replace("STATE_", "") for c in simulation_cols]
+    state_portion = (curr_game
+        .loc[:, simulation_cols]
+        .copy()
         .reset_index(drop=True)
-        .drop(columns=['game_seconds_remaining']))
-    cols = np.array(diffs.columns)
-    inds = np.argmax(diffs.to_numpy(), axis=1)
-    events = cols[inds]
-    if len(events) >= 3:
-        prev3, prev2, prev1 = events[-3], events[-2], events[-1]
-    elif len(events) == 2:
-        prev3, prev2, prev1 = '#', events[-2], events[-1]
-    elif len(events) == 1:
-        prev3, prev2, prev1 = '#', '#', events[-1]
-    else:
-        prev3, prev2, prev1 = '#', '#', '#'
[…]

### 2024-03-29 Latex/draft.tex

@@
 separate XGBoost models for each team to learn their unique patterns 
-and strategies. Leveraging the {\tt .predict\_proba()} method of XGBoost, we generated 
-probabilistic predictions at various stages of a game, allowing us to 
-plot the evolving probability of each team winning throughout the match
-(see~\ref{fig:xgboost_res}). 
-This approach provided insights into momentum shifts (when one team 
-has an advantage due to them having more players on the ice due to 
-a penalty by the opposing team), key moments (such as goals and 
-penalties, and critical plays influencing game outcomes).
-
-% Images
-\begin{figure}[H]
-    \centering
-    \subfigure[]{\includegraphics[width=0.6\textwidth]{images/good_xgb_example.png}} 
-    \subfigure[]{\includegraphics[width=0.6\textwidth]{images/good_xgb_example(2).png}} 
-    \caption{For both of these plots, the home team is above the the 50\% line, with the
-    away team below. The title has the final score of the game. \\
-        \textbf{(a)} The live win probabilities for the Nashville Predators at the Vegas Golden Knights. 
-    Notice the probability shifts after one team scores a goal.  \\
-        \textbf{(b)} The live win probabilities for the Calgary Flames at the Winnipeg Jets.
-        Notice the stark changes in probabilities after each goal. Also notice the
-        shift in probabilities even when no goals are scored. This suggests
-        the XGBoost model is able to pick up on momentum shifts.}
-    \label{fig:xgboost_res}
-\end{figure}
+and strategies. We chose to create an XGBoost model for each team
+because we felt it would be more accurate than a single model for all teams.
+This is because some teams do very well and some teams do very poorly, and
+we felt this difference necessitated a model for each team.
 
@@
 
+\subsection{XGBoost}
+
+As mentioned in the methods section, we trained an XGBoost model for each team in the NHL. One main benefit of this
+approach is the ability to leverage the {\tt .predict\_proba()} method of XGBoost.
+Using this method, we generated 
+probabilistic predictions at various stages of a game, allowing us to 
+plot the evolving probability of each team winning throughout the match
+(see~\ref{fig:xgboost_res}). 
+This approach provided insights into momentum shifts (when one team 
+has an advantage due to them having more players on the ice due to 
+a penalty by the opposing team), key moments (such as goals and 
+penalties, and critical plays influencing game outcomes).
+
+\begin{figure}[H]
+    \centering
+    \subfigure[]{\includegraphics[width=0.6\textwidth]{images/good_xgb_example.png}} 
+    \subfigure[]{\includegraphics[width=0.6\textwidth]{images/good_xgb_example(2).png}} 
+    \caption{For both of these plots, the home team is above the the 50\% line, with the
+    away team below. The title has the final score of the game. \\
+        \textbf{(a)} The live win probabilities for the Nashville Predators at the Vegas Golden Knights. 
+    Notice the probability shifts after one team scores a goal.  \\
+        \textbf{(b)} The live win probabilities for the Calgary Flames at the Winnipeg Jets.
+        Notice the stark changes in probabilities after each goal. Also notice the
+        shift in probabilities even when no goals are scored. This suggests
+        the XGBoost model is able to pick up on momentum shifts.}
+    \label{fig:xgboost_res}
+\end{figure}
+
 \section{Ethical Considerations}

### 2024-03-30 Latex/draft.tex
finished analysis and results for MCMC and accuracy

@@
 \begin{abstract}
-    The goal of this project is simple: predict the outcomes of NHL games from any given state.
-    As simple as the problem statement is, however, the solution is not so straightforward.
-    To solve this problem, we will use a variety of machine learning techniques, including logistic regression,
-    XGBoost, and ARIMA models. Additionally, we utilize a form of MCMC to simulate the outcomes of games from any given
-    state. Our hypothesis is that we will be able to successfully predict the outcomes of NHL games with a high degree of accuracy
-    using these tools.
+    The goal of this project is simple: predict the outcomes of NHL games from any given state during the game.
+To solve this problem we use three primary methods: Bayesian Network to train a DAG, XGBoost on game states, and an MCMC game simulator.
+We use each of these methods to generate win probabilities for each time step throughout various NHL games, thus emulating a live win probability.
+Finally, we analyze the accuracy of these three methods and considered ethical implications of our results.
 \end{abstract}
@@
 
-To simulate a game, we performed this Monte Carlo algorithm that acted like a random walk through the Markov Chain:
+To simulate a game, we performed this Monte Carlo algorithm that acted like a random walk through the Markov Chain (Note that the KDE\_times() is a KDE model fit on the amount of seconds that transpired between
+each hockey event for all NHL games over the course of 13 years):
 \begin{algorithm}[H]
@@
         \State $s_2 \gets "\#"$
-        %\State $\text{event\_counts} \gets \text{dictionary()}$
+        \State $\text{event\_counts} \gets \text{dictionary()}$
 
@@
             \State $\text{next\_state} \gets \text{sample from } \text{MC(curr\_state)}$
-            %\State $\text{event\_counts}[\text{next\_state}] \mathrel{+}= 1$
+            \State $\text{event\_counts}[\text{next\_state}] \mathrel{+}= 1$
 
@@
 \end{algorithm}
-The KDE\_times() is a KDE model fit on the amount of seconds that transpired between 
-each hockey event for all NHL games over the course of 13 years. This KDE model 
-resembled a poisson distribution with $\lambda \approx 18$. During the course of 
-the algorithm, we maintained a count of all of the events that transpired. Upon 
-termination of the simulation, we compared home goals to away goals and termed 
-a winner. If there was a tie, we would run the algorithm until another team scored0 
-a goal, thus simulating overtime per NHL rules.
 
-To predict a probability of a certain team winning a hockey game, we would run 
-50 simulations with our algorithm, but we would initialize the starting states 
-$(s_0, s_1, s_2)$ to be the most current events in the hockey game. Out of the 
-50 simulations, we would compute 
-$P(\text{home winning}) = \{\text{\# home simulation wins}\}/50$ and 
+To predict a team's winning probability, we would simulate 50 hockey games with the initial starting states $(s_0, s_1, s_2)$ set to the most current events in the hockey game.
+By looking at each games
+final event counts, we could compute each game's winner. Then we computed the 
+probability as $P(\text{home winning}) = \{\text{\# home simulation wins}\}/50$ and 
 $P(\text{away winning}) = 1-P(\text{home winning})$.
-
-To evaluate our simulation's effectiveness at accurately modeling actual hockey 
-games, we combined a dataset with the final event counts for 1500 actual hockey 
-games and for 1500 simulated games. We then performed PCA, t-SNE and UMAP 
-dimensionality reductions using various perplexity and neighbor hyperparameters 
-to see if these algorithms clustered the synthetic games and actual games in 
-different clusters. As shown below in~\ref{fig:simulation_v_actual}, the simulated
-games and actual games are all clustered together, thus demonstrating that our
-simulation effectively emulates live NHL games.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.9\textwidth]{images/pca_tsne_umap_sim_act.png}
-    \caption{PCA, t-SNE, and UMAP performed on a dataset with final event counts for 1500 simulated and 1289 actual NHL games.
-The joined grouping demonstrates that our simulation accurately emulates live NHL games}
-    \label{fig:simulation_v_actual}
-\end{figure}
 
@@
 
+\subsection{MCMC Simulator}
+
+To evaluate our simulation's effectiveness at accurately modeling actual hockey 
+games, we combined a dataset with the final event counts for 1500 actual hockey 
+games and for 1500 simulated games. We then performed two experiments. First, We 
+trained a KMeans, Logistic Regression, RandomForest and XGBoost model on this 
+dataset with default parameters, to see 
+if they could classify games as simulated or real. With a 70-30 train-test split, 
+the accuracies on the test set were 47.9\%, 48.5\%, 67.5\% and 69.3\% respectively. 
+Second, we fit and transformed this dataset using PCA, t-SNE and UMAP 
+dimensionality reductions using various perplexity and neighbor hyperparameters to see 
+if these algorithms clustered the synthetic games and actual games in 
+different clusters. As shown below in ~\ref{fig:simulation_v_actual}, the simulated
+games and actual games are all clustered together. From these two experiments, we conclude
+that our simulator effectively models live NHL games.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.9\textwidth]{images/pca_tsne_umap_sim_act.png}
+    \caption{PCA, t-SNE, and UMAP performed on a dataset with final event counts for 1500 simulated and 1500 actual NHL games.
+The joined grouping demonstrates that our simulation accurately emulates live NHL games.
+Note that we performed the t-SNE and UMAP on many different values of perplexity and neighbors, but all of the plots looked the same, so we just show this one}
+    \label{fig:simulation_v_actual}
+\end{figure}
+
+\subsection{Method Comparison}
+Each of the methods described above had one general purpose: accurately predict the outcome of a hockey game regardless of the point in the game.
+To evaluate our models, we randomly selected 50 games outside of the training data, and for various time steps throughout these games we had each model calculate a win probability.
+We then used the probability threshold of 0.5 to determine whether or not each model predicted if the home team won or lost.
+We then computed each model's accuracy across all 50 games for each of the time steps.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.9\textwidth]{images/game_accuracies.png}
+    \caption{Each of our model's predictive accuracy over the course of 50 games}
+    \label{fig:game_accuracies}
+\end{figure}
+
+In figure ~\ref{fig:game_accuracies} we note that each model increased in accuracy as the game progressed.
+We also notice that the MCMC simulation method appears to have the highest accuracy regardless of the point in the game while XGBoost appears to be the least accurate.
+We infer that the XGBoost models were overfit, since they appeared to perform very well on games within the training set.
+Also, during experimentation with the models, we noticed that the MCMC simulator and the Bayesian Network tended to produce higher probabilities for the away team winning, and the XGBoost model tended to give higher probabilities to the home team winning.
+The MCMC and Bayesian Network are structured to model the flow of events, so it appears that they picked up on a more predictive flow in how away teams play.
+
 \section{Ethical Considerations}

### 2024-03-30 Latex/draft.tex
final for draft

@@
         Notice the stark changes in probabilities after each goal. Also notice the
-        shift in probabilities even when no goals are scored. This suggests
-        the XGBoost model is able to pick up on momentum shifts.}
+        shift in probabilities even when no goals are scored.}
     \label{fig:xgboost_res}
@@
 
-\section{Conclusions}
-% Your content for this section goes here
+\section{Conclusion}
+
+In this project, we used three different methods to predict the outcomes of NHL games.
+We used a Bayesian Network to model the conditional dependencies of key features, XGBoost to model the play-by-play dynamics of each team, and an MCMC simulator to simulate the flow of events in a hockey game.
+We found that the MCMC simulator was the most accurate, with the Bayesian Network coming in second, and XGBoost coming in last.
+The win percentages predicted are more for entertainment and educational purposes, and should not be used for gambling or performance purposes, as they do not have a incredibly high degree of accuracy.
+There are many variables in a hockey game that affect win probability, and given more time and more data we would love to model other factors such as momentum, historical success, coaching, etc.
+Our project was limited by the data
+we had access to and computational resources, but nonetheless we were able to produce a model that can predict the outcome of NHL games with a reasonable degree of accuracy.
+
+% new page
+\newpage
 

### 2024-04-10 Latex/draft.tex

@@
 nine events: blocked shot, faceoff, giveaway, goal, hit, missed shot, penalty, shot, and takeaway. These events are attributed to the
-team and player that performs the event. We only take into consideration the team that performs the event and discard the player information.
+team and player that perform the event. We only take into consideration the team that performs the event and discard the player information.
 
@@
 \section{Ethical Considerations}
-Predicting win percentages or outcomes in hockey games, like any sport, raises several ethical considerations. Here are some key points we want to address:
+Predicting win percentages or outcomes in hockey games, like any sport, raises several ethical considerations relating to the fariness and integrity of the sport.
+We believe that the main ethical issue with our model is the potential for it to be used for gambling purposes.
+We do not condone the use of our model for gambling, and the purpose of our model is purely for educational and entertainment purposes.
+However, we recognize that the primary reason the the public uses sports predicting models is for gambling reasons.
 
-\begin{itemize}[label=\textbullet]
-\item \textbf{Gambling and Addiction}:
-Our win percentages and predictions might be used by those who wish to gamble, which could lead to addiction and financial harm, especially if undue trust in placed in these methods.
-Any publication of these methods or predictions would be accompanied by promoting healthy and responsible gambling practices.
+The juxtaposition of our reasoning for completing this research and how we believe this research would be used is an interesting ethical dilemma. 
+One could argue that because our model will be used for gambling, we should not have released or published this work in the first place. 
+However, we believe that the technological and mathematical advancements that we have made in this project are valuable and should be shared with the public. 
+We also believe that the potential for our model to be used for gambling is not a reason to withhold our work from the public.
 
-\item \textbf{Fairness and Integrity of the Game}:
-Sometimes, coaches and players becoming aware of their chance of winning can affect how the game is played, potentially harming the integrity of the sport.
-We must be careful to not provide an unfair advantage to any team or player.
-Inaccurate predictions could lead a team to believe that the game is out of reach when it isn't, and we want to avoid that.
+Gambling is harmful because it can lead to addiction, financial ruin, and other negative consequences.
+Gambling can also affect outcomes of games unfairly if players, coaches, owners, or referees illegally participate in gambling.
+Corrupt actors could create a destructive feedback loop using our predictions to gamble and then therefore influence the outcome of the game to match our predictions.
+We wish to make it clear that we are vehemently against sports gambling, and we support laws, policies, and practices that prevent gambling from negatively affecting sports and encourage healthy and responsible gambling practices.
 
-\end{itemize}
-
-Overall, our predictions, like any, should not be considered declarative for gambling or performance purposes, but are rather an interesting exposition into the complexity of sporting events.
+Additionally, there are no privacy concerns with our data as we only collected publicly availalbe NHL data.
 

### 2024-04-10 Latex/draft.tex
Made latex changes (no benj changes yet).

@@
 events that occurred. For example, if the home team won a faceoff, the home team lost the puck, and 
-then the away team shot the puck and missed, the markov chain would look like figure~\ref{fig:markov_chain_sample}.
+then the away team shot the puck and missed, the markov chain would look like Figure~\ref{fig:markov_chain_sample}.
 
-\begin{figure}[H]
+\begin{figure}
     \centering
-    \includegraphics[width=0.8\textwidth]{images/markov_chain_sample.png}
-    \caption{Sample Markov Chain for the current state (home team won faceoff, 
-    home team lost puck, away team missed shot) with the probabilities of the next event 
-    sorted from highest to lowest}
+    \begin{tabular}{ |c|c| } 
+    \hline
+        \multicolumn{2}{|c|}{\textbf{Sample Markov Chain}} \\
+     \hline
+     \textit{Next Event} & \textit{Transition Probability} \\ 
+     \hline
+     Shot Away & 0.1853207 \\ 
+     \hline
+     Blocked Shot Away & 0.1138666 \\ 
+     \hline
+     \vdots & \vdots \\
+    \hline
+    Penalty Away & 0.0167210 \\
+    \hline
+    Goal Home & 0.0085793 \\
+    \hline
+    \end{tabular}
+    \caption{The transition events and probabilities when the previous three events were Faceoff Home, Giveaway Home, and Missed
+    Shot Away, respectively.}
     \label{fig:markov_chain_sample}
 \end{figure}
+
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=0.8\textwidth]{images/markov_chain_sample.png}
+%     \caption{Sample Markov Chain for the current state (home team won faceoff, 
+%     home team lost puck, away team missed shot) with the probabilities of the next event 
+%     sorted from highest to lowest}
+%     % \label{fig:markov_chain_sample}
+% \end{figure}
 
@@
 \end{align*}
-Where $A,B,C,D$ represent events that can occur in a game, and the tuple (A,B,C,D) represents that "A then 
-B then C then D" happened right after each other in a game.
+Where $A,B,C,D$ represent events that can occur in a game, and the tuple (A,B,C,D) represents that ``A then 
+B then C then D'' happened right after each other in a game.
 
-To simulate a game, we performed this Monte Carlo algorithm that acted like a random walk through the Markov Chain (Note that the KDE\_times() is a KDE model fit on the amount of seconds that transpired between
-each hockey event for all NHL games over the course of 13 years):
-\begin{algorithm}[H]
+To simulate a game, we performed this Monte Carlo Algorithm (see Algorithm~\ref{alg:monte_carlo_algorithm}) that acted like a random walk through the Markov Chain (Note that the {\tt KDE\_times()} is a KDE model fit on the amount of seconds that transpired between
+each hockey event for all NHL games over the course of 13 years).
+\begin{algorithm}
     \caption{Simulation Algorithm}
@@
 of factors other than goals that could affect the win probabilities, so the predicted probabilities are little more than an over probability calculation of goals and time reamining given historical data.
-We see this in the probability graph below, where the changes in probability correspond to the goals scored in the game. A sample win probability graph is seen below in ~\ref{fig:bayesian_graph}
+We see this in the probability graph below, where the changes in probability correspond to the goals scored in the game.
+A sample win probability graph is seen below in Figure~\ref{fig:bayesian_graph}.
 
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 plot the evolving probability of each team winning throughout the match
-(see~\ref{fig:xgboost_res}). 
+(see Figure~\ref{fig:xgboost_res}). 
 This approach provided insights into momentum shifts (when one team 
@@
 
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 if these algorithms clustered the synthetic games and actual games in 
-different clusters. As shown below in ~\ref{fig:simulation_v_actual}, the simulated
+different clusters. As shown below in Figure~\ref{fig:simulation_v_actual}, the simulated
 games and actual games are all clustered together. From these two experiments, we conclude
@@
 Each of the methods described above had one general purpose: accurately predict the outcome of a hockey game regardless of the point in the game.
-To evaluate our models, we randomly selected 50 games outside of the training data, and for various time steps throughout these games we had each model calculate a win probability.
+To evaluate our models, we randomly selected 100 games outside of the training data, and for various time steps throughout these games we had each model calculate a win probability.
 We then used the probability threshold of 0.5 to determine whether or not each model predicted if the home team won or lost.
@@
 
-\begin{figure}[H]
+\begin{figure}
     \centering
-    \includegraphics[width=0.9\textwidth]{images/game_accuracies.png}
-    \caption{Each of our model's predictive accuracy over the course of 50 games}
+    \begin{tabular}{ |c|c|c|c| } 
+    \hline
+        \multicolumn{4}{|c|}{\textbf{Accuracy Comparison Between Models}} \\
+     \hline
+     \textit{Game Seconds Played} & \textit{MCMC} & \textit{XGBoost} & \textit{Bayesian} \\ 
+     \hline
+     0 & a & a & a \\ 
+     \hline
+     600 & a & a & a \\ 
+     \hline
+     1200 & a & a & a \\
+    \hline
+    1800 & a & a & a \\
+    \hline
+    2400 & a & a & a \\
+    \hline
+    3000 & a & a & a \\
+    \hline
+    3600 & a & a & a \\
+    \hline
+    \end{tabular}
+    \caption{Each of our model's predictive accuracy over the course of 100 games}
     \label{fig:game_accuracies}
@@
 
-In figure ~\ref{fig:game_accuracies} we note that each model increased in accuracy as the game progressed.
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=0.9\textwidth]{images/game_accuracies.png}
+%     \caption{Each of our model's predictive accuracy over the course of 50 games}
+%     \label{fig:game_accuracies}
+% \end{figure}
+
+In Figure~\ref{fig:game_accuracies} we note that each model increased in accuracy as the game progressed.
 We also notice that the MCMC simulation method appears to have the highest accuracy regardless of the point in the game while XGBoost appears to be the least accurate.

### 2024-04-10 Latex/draft.tex
Change tables from 'figure' to 'table.

@@
 events that occurred. For example, if the home team won a faceoff, the home team lost the puck, and 
-then the away team shot the puck and missed, the markov chain would look like Figure~\ref{fig:markov_chain_sample}.
+then the away team shot the puck and missed, the markov chain would look like Table~\ref{tab:markov_chain_sample}.
 
-\begin{figure}
+\begin{table}
     \centering
@@
     Shot Away, respectively.}
-    \label{fig:markov_chain_sample}
-\end{figure}
+    \label{tab:markov_chain_sample}
+\end{table}
 
@@
 
-\begin{figure}
+\begin{table}
     \centering
@@
     \caption{Each of our model's predictive accuracy over the course of 100 games}
-    \label{fig:game_accuracies}
-\end{figure}
+    \label{tab:game_accuracies}
+\end{table}
 
@@
 
-In Figure~\ref{fig:game_accuracies} we note that each model increased in accuracy as the game progressed.
+In Figure~\ref{tab:game_accuracies} we note that each model increased in accuracy as the game progressed.
 We also notice that the MCMC simulation method appears to have the highest accuracy regardless of the point in the game while XGBoost appears to be the least accurate.

### 2024-04-12 jeff_py.py
syncing

@@ -1,0 +1,50 @@
+import multiprocessing
+import time
+from tqdm.auto import tqdm
+import random
+
+total_processed = 0
+start_time = time.time()
+game_ids = [1,2,3,4,5]
+
+num_processes = 2
+
+def update_progress(result, total_processed=[0]):
+    global start_time, game_ids
+    elapsed_time = time.time() - start_time
+    total_processed[0] += num_processes
+    remaining_games = len(game_ids) - total_processed[0]
+    estimated_time_left = elapsed_time / total_processed[0] * remaining_games
+    print(elapsed_time, total_processed[0], remaining_games, estimated_time_left)
+    print(f"Game {result} completed. Estimated time remaining: {estimated_time_left:.2f} seconds")
+
+def get_game_probabilities(game_id):
+    start_time = time.time()
+    # Simulate game and get probabilities
+    # Replace this with your actual logic (should take some time)
+    pbar = tqdm(total=10)
+    pbar.set_description(f"Processing game {game_id}")
+    for i in range(10):
+        time.sleep(random.random()*2)
+        pbar.update(1)
+    elapsed_time = time.time() - start_time
+    print(f"Game {game_id} probabilities calculated in {elapsed_time:.2f} seconds")
+    update_progress(game_id)
+
+def simulate_games_multi():
+    global num_processes
+    game_ids = [1, 2, 3, 4, 5]  # Replace with your game IDs
+
+
+    # Create a pool of processes
+    pool = multiprocessing.Pool(processes=num_processes)
+
+    # Use map with a callback for progress updates
+    results = pool.map(get_game_probabilities, game_ids)
+
+    # Close the pool after use
+    pool.close()
+    pool.join()
+
+if __name__ == "__main__":
+    simulate_games_multi()

### 2024-04-12 simulate_games.py

@@ -1,3 +1,6 @@
+import multiprocessing
+import concurrent
+
 import pandas as pd
 import pickle
 from tqdm import tqdm
@@ -5,9 +8,11 @@ import numpy as np
 import os
 import sys
 import time
-from jeffutils.utils import stack_trace
+from jeffutils.utils import stack_trace, str_to_list_py
 
 from py_files.simulator import get_simulation_prob
+
+start_time = time.time()
 
 def simulate_games(n_games:int, save_path:str):
     
@@ -125,7 +130,11 @@ def simulate_games(n_games:int, save_path:str):
 
     return games_full
 
-def get_game_probabilities(game_id):
+def get_game_probabilities(inpt):
+    
+    start = time.time()
+    
+    game_id, game_index_i, n_games = inpt
     
     df_full = pd.read_feather("data/play_by_play/play_by_play_full_state_space.feather")
     
@@ -163,7 +172,7 @@ def get_game_probabilities(game_id):
     for i, ind in enumerate(rows_to_simulate):
         inds = np.arange(max(0, ind-4), ind+1)
         curr_snapshot = state_portion.loc[inds, :].copy()
-        print("Seconds remaining:", curr_snapshot['game_seconds_remaining'].values[-1], f"({i+1}/{len(rows_to_simulate)})")
+        print(f"{game_id} ({game_index_i}/{n_games}) Seconds remaining:", curr_snapshot['game_seconds_remaining'].values[-1], f"({i+1}/{len(rows_to_simulate)})")
         
         home_prob, away_prob = get_simulation_prob(curr_snapshot, prob_mc, kde_seconds, verbose=True)
         simulation_probabilities.append((home_prob, away_prob))
@@ -177,6 +186,11 @@ def get_game_probabilities(game_id):
     
     with open(f"data/pickles/simulation_probs_{game_id}.pickle", "wb") as f:
         pickle.dump(save_stuff, f)
+        
+    global start_time
+    minutes_elapsed = (time.time() - start_time) / 60
+    curr_time = (time.time() - start) / 60
+    print(f"\n{game_index_i+1}/{n_games} total_min:{round(minutes_elapsed, 2)} curr_min:{round(curr_time, 2)}\n")
         
 def simulate_games():
     game_ids = [2022020839, 2023020412, 2022020727, 2022030154, 
@@ -193,8 +207,13 @@ def simulate_games():
                 2022020753, 2023020475, 2023020192, 2022020662, 
[…]

### 2024-04-12 Latex/draft.tex
added the accuracy table

@@
 nine events: blocked shot, faceoff, giveaway, goal, hit, missed shot, penalty, shot, and takeaway. These events are attributed to the
-team and player that perform the event. We only take into consideration the team that performs the event and discard the player information.
+team and player that perform the event. We only take into consideration the team that performs the event and discard the player information to reduce noise and to avoid many one-hot encoded columns.
 
@@
 \subsection{Method Comparison}
-Each of the methods described above had one general purpose: accurately predict the outcome of a hockey game regardless of the point in the game.
-To evaluate our models, we randomly selected 100 games outside of the training data, and for various time steps throughout these games we had each model calculate a win probability.
-We then used the probability threshold of 0.5 to determine whether or not each model predicted if the home team won or lost.
-We then computed each model's accuracy across all 50 games for each of the time steps.
+Each of our methods had one general purpose: predict the winner of the game regardless of the point of the game.
+To evaluate our models, we randomly selected 100 games outside of the training data.
+To ensure that the test set represented the actual data, we selected the test set with 50 home wins and 50 away wins, and we also made sure there was an even representation of each of the 32 NHL teams.
+We then used each model to predict a winner at various points throughout each game.
+From these predictions on the 100 games, we computed each model's accuracy for each point in the game.
 
@@
     \centering
-    \begin{tabular}{ |c|c|c|c| } 
+    \begin{tabular}{ |c|c|c|c|c| } 
     \hline
-        \multicolumn{4}{|c|}{\textbf{Accuracy Comparison Between Models}} \\
+    \textbf{Game Seconds} & \textbf{Game State} & \textbf{MCMC} & \textbf{XGBoost} & \textbf{Bayesian} \\ 
     \hline
-     \textit{Game Seconds Played} & \textit{MCMC} & \textit{XGBoost} & \textit{Bayesian} \\ 
+    0 & Q1 & 47\% & 56\% & 51\% \\ 
     \hline
-     0 & a & a & a \\ 
+    900 & Q2 & 68\% & 55\% & 69\% \\
     \hline
-     600 & a & a & a \\ 
+    1800 & Q3 & 73\% & 58\% & 69\% \\
     \hline
-     1200 & a & a & a \\
+    2700 & Q4 & 79\% & 77\% & 80\% \\
     \hline
-    1800 & a & a & a \\
+    3540 & Last min of Q4 & 94\% & 86\% & 91\% \\
     \hline
-    2400 & a & a & a \\
-    \hline
-    3000 & a & a & a \\
-    \hline
-    3600 & a & a & a \\
+     \(\sim \)3800 & OT & 96\% & 93\% & 100\% \\
     \hline
     \end{tabular}
-    \caption{Each of our model's predictive accuracy over the course of 100 games}
+    \caption{Results of each model's accuracy at various points throughout 100 games sampled from the test set}
     \label{tab:game_accuracies}
@@
 In Figure~\ref{tab:game_accuracies} we note that each model increased in accuracy as the game progressed.
-We also notice that the MCMC simulation method appears to have the highest accuracy regardless of the point in the game while XGBoost appears to be the least accurate.
+We also notice that the MCMC simulation method appears to have the highest accuracy after the start of the game while XGBoost appears to be the least accurate.
 We infer that the XGBoost models were overfit, since they appeared to perform very well on games within the training set.
-Also, during experimentation with the models, we noticed that the MCMC simulator and the Bayesian Network tended to produce higher probabilities for the away team winning, and the XGBoost model tended to give higher probabilities to the home team winning.
-The MCMC and Bayesian Network are structured to model the flow of events, so it appears that they picked up on a more predictive flow in how away teams play.
+Also, during experimentation with the models, we noticed that the MCMC simulator tended to produce higher probabilities for the away team winning, and the XGBoost model tended to give higher probabilities to the home team winning.
+The Bayesian Network was unbiased towards the home or away team.
+We trained the MCMC model on the flow of events in the game, so it appears that the model picked up on a more predictive flow in how away teams play.
+Though the MCMC model was generally the most accurate, it takes about 2 minutes to simulate 50 games and get a single prediction while XGBoost and the Bayesian Network can make a prediction in under two seconds.
 

### 2024-04-12 Latex/draft.tex
Moved algorithm 1 to the appendix (save space).

@@
 
-To simulate a game, we performed this Monte Carlo Algorithm (see Algorithm~\ref{alg:monte_carlo_algorithm}) that acted like a random walk through the Markov Chain (Note that the {\tt KDE\_times()} is a KDE model fit on the amount of seconds that transpired between
+To simulate a game, we performed this Monte Carlo Algorithm (see Algorithm~\ref{alg:monte_carlo_algorithm} in Appendix~\ref{mcmc_app}) that acted like a random walk through the Markov Chain (Note that the {\tt KDE\_times()} is a KDE model fit on the amount of seconds that transpired between
 each hockey event for all NHL games over the course of 13 years).
-\begin{algorithm}
-    \caption{Simulation Algorithm}
-    \begin{algorithmic}[1]
-        \State $\text{time\_remaining} \gets 3600$
-        \Comment{NHL games are 3600 seconds}
-        \State $s_0 \gets "\#"$
-        \Comment{The "\#" symbol represents the start of a game}
-        \State $s_1 \gets "\#"$
-        \State $s_2 \gets "\#"$
-        \State $\text{event\_counts} \gets \text{dictionary()}$
+% \begin{algorithm}
+%     \caption{Simulation Algorithm}
+%     \begin{algorithmic}[1]
+%         \State $\text{time\_remaining} \gets 3600$
+%         \Comment{NHL games are 3600 seconds}
+%         \State $s_0 \gets "\#"$
+%         \Comment{The "\#" symbol represents the start of a game}
+%         \State $s_1 \gets "\#"$
+%         \State $s_2 \gets "\#"$
+%         \State $\text{event\_counts} \gets \text{dictionary()}$
 
-        \While{$\text{time\_remaining} > 0$}
-            \State $\text{next\_state} \gets \text{sample from } \text{MC(curr\_state)}$
-            \State $\text{event\_counts}[\text{next\_state}] \mathrel{+}= 1$
+%         \While{$\text{time\_remaining} > 0$}
+%             \State $\text{next\_state} \gets \text{sample from } \text{MC(curr\_state)}$
+%             \State $\text{event\_counts}[\text{next\_state}] \mathrel{+}= 1$
 
-            \State $s_0 \gets s_1$
-            \State $s_1 \gets s_2$
-            \State $s_2 \gets \text{next\_state}$
+%             \State $s_0 \gets s_1$
+%             \State $s_1 \gets s_2$
+%             \State $s_2 \gets \text{next\_state}$
 
-            \State $\text{event\_time} \gets \text{sample from } \text{KDE\_times()}$
-            \State $\text{time\_remaining} \gets (\text{time\_remaining} - \text{event\_time})$
-        \EndWhile
-    \end{algorithmic}
-    \label{alg:monte_carlo_algorithm}
-\end{algorithm}
+%             \State $\text{event\_time} \gets \text{sample from } \text{KDE\_times()}$
+%             \State $\text{time\_remaining} \gets (\text{time\_remaining} - \text{event\_time})$
+%         \EndWhile
+%     \end{algorithmic}
+%     \label{alg:monte_carlo_algorithm}
+% \end{algorithm}
 
@@
 
+\newpage
+
+\section*{Appendix}
+
+\appendix
+\section{MCMC Sampling Algorithm}
+\label{mcmc_app}
+
+\begin{algorithm}
+    \caption{Simulation Algorithm}
+    \begin{algorithmic}[1]
+        \State $\text{time\_remaining} \gets 3600$
+        \Comment{NHL games are 3600 seconds}
+        \State $s_0 \gets "\#"$
+        \Comment{The "\#" symbol represents the start of a game}
+        \State $s_1 \gets "\#"$
+        \State $s_2 \gets "\#"$
+        \State $\text{event\_counts} \gets \text{dictionary()}$
+
+        \While{$\text{time\_remaining} > 0$}
+            \State $\text{next\_state} \gets \text{sample from } \text{MC(curr\_state)}$
+            \State $\text{event\_counts}[\text{next\_state}] \mathrel{+}= 1$
+
+            \State $s_0 \gets s_1$
+            \State $s_1 \gets s_2$
+            \State $s_2 \gets \text{next\_state}$
+
+            \State $\text{event\_time} \gets \text{sample from } \text{KDE\_times()}$
+            \State $\text{time\_remaining} \gets (\text{time\_remaining} - \text{event\_time})$
+        \EndWhile
+    \end{algorithmic}
+    \label{alg:monte_carlo_algorithm}
+\end{algorithm}
+
 \end{document}

### 2024-04-12 Latex/draft.tex
Pushing changes.

@@
 
-Hockey, however, is a different story. While there are some methods used to predict the outcome of National Hockey League (NHL) games, these models
+Hockey; however, is a different story. While there are some methods used to predict the outcome of National Hockey League (NHL) games, these models
 typically belong to sport books and their nuances are not publicly disclosed. Additionally, hockey analytics is not as
-developed as it is in other sports, such as basketball or baseball. This lack of model transparency and public interest in hockey analytics
+developed as it is in other sports, such as basketball or baseball~\cite{Brooks-Davis}. This lack of model transparency and public interest in hockey analytics
 makes predicting the outcomes of NHL games a very underdeveloped and challenging problem. Previous attempts and research into predicting NHL games
@@
 
-In addition to model research, some research has also gone into developing new features that can be used to better predict the game outcomes. 
+In addition to model research, some effort has also gone into developing new features that can be used to better predict the game outcomes. 
 The two biggest engineered classes of features are the Corsi
@@
  We were unable to locate the original blog posts talking about these metrics, but a good article to learn more about the math can be
- found here \url{https://thehockeywriters.com/corsi-fenwick-stats-what-are-they/}.} metrics (both around 2007).
+ found here \url{https://thehockeywriters.com/corsi-fenwick-stats-what-are-they/}.} metrics (both around 2007); however, we do not use these
+ in our approach because future research shows they do not improve model performance~\cite{rinkstats_corsi_fenwick}.
 
 Our project seeks a similar outcome to the research mentioned above: predict the outcomes of NHL games. Not only this,
-but we seek to provide live, acturate win probabilities for any given game state. Despite the simplicity of the problem statement, 
-as mentioned, the solution is not so straightforward. The NHL provides fast-paced games with many events
+but we seek to provide live, acturate, win probabilities for any given game state. Despite the simplicity of the problem statement, 
+the solution is not so straightforward. The NHL provides fast-paced games with many events
 occuring in quick succession. Our goal is to use this abundance of data and new approaches to build upon previous research.
@@
 Each game in a season is given a unique identifier ({\tt game\_id}), which is constant across all events in a game. Every event that occurs in a game
-will be stored in the {\tt event\_type} column. There are 17 unique event types, including things such as game start, faceoff, shot, hit, and goal.
+will be stored in the {\tt event\_type} column. There are 17 unique event types, including game start, faceoff, shot, hit, and goal.
 Most of these event types are not relevant to our analysis, so we remove them from the dataset. After removing the unnecessary events, we are left with
-nine events: blocked shot, faceoff, giveaway, goal, hit, missed shot, penalty, shot, and takeaway. These events are attributed to the
+nine: blocked shot, faceoff, giveaway, goal, hit, missed shot, penalty, shot, and takeaway. These events are attributed to the
 team and player that perform the event. We only take into consideration the team that performs the event and discard the player information to reduce noise and to avoid many one-hot encoded columns.
@@
 
-We found that our data did not contain any missing values that was not easily explainable. For example, if a game is starting, there will be no
+We found that our data did not contain any missing values that were not easily explainable. For example, if a game is starting, there will be no
 events for penalties, which will result in a {\tt NaN} value in the penalties column. Additionally, any data that was confusing or not easily explainable
@@
 \subsection{Bayesian Network}
-We first used a Bayesian Network to establish a benchmark for probability using several key features.
+We first used a Bayesian network to establish a benchmark for probability using several key features.
 
@@
 The computational complexity of Bayesian Network inference is high, with exact inference being an NP-hard problem~\cite{pmlr-vR0-chickering95a}. 
-Using the python package pgmpy, we originally tried to fit a network with all 26 of our features, but our computational resources failed to fit this network.
+Using the python package {\tt pgmpy}, we originally tried to fit a network with all 26 of our features, but our computational resources failed to fit this network.
 Then, to get a baseline for our future predictions, we simply fitted the model with the base features of time remaining (tr), home goals (hg), away goals (ag), home shots (hs), away shots (as), home blocked shots (hbs), and away blocked shots (abs) in order to predict wins (w).
@@
 
+XGBoost was chosen because of its ability to handle large datasets and model complex relationships between features. 
+XGBoost is able to attain a high degree of accuracy with a relatively small training time or overhead and was an ideal choice for our research.
+
 \subsection{MCMC Game Simulation}
@@
 To simulate hockey games, we created a Markov Chain where the states are a tuple of three consecutive
-events that occurred. For example, if the home team won a faceoff, the home team lost the puck, and 
+events that occurred. For example, if the home team won a faceoff, lost the puck, and 
 then the away team shot the puck and missed, the markov chain would look like Table~\ref{tab:markov_chain_sample}.
@@
 
-To simulate a game, we performed this Monte Carlo Algorithm (see Algorithm~\ref{alg:monte_carlo_algorithm} in Appendix~\ref{mcmc_app}) that acted like a random walk through the Markov Chain (Note that the {\tt KDE\_times()} is a KDE model fit on the amount of seconds that transpired between
+To simulate a game, we performed this Monte Carlo Algorithm (see Algorithm~\ref{alg:monte_carlo_algorithm} in Appendix~\ref{mcmc_app})
+that acts as a random walk through the Markov Chain (Note that the {\tt KDE\_times()} is a KDE model fit on the amount of seconds that transpired between 
 each hockey event for all NHL games over the course of 13 years).
@@
 
-To predict a team's winning probability, we would simulate 50 hockey games with the initial starting states $(s_0, s_1, s_2)$ set to the most current events in the hockey game.
-By looking at each games
-final event counts, we could compute each game's winner. Then we computed the 
+To predict a team's winning probability, we would simulate 50 hockey games with the initial starting states 
+$(s_0, s_1, s_2)$ set to the most recent events in the hockey game. By looking at each games
+final event counts we compute the winner. We then compute the 
 probability as $P(\text{home winning}) = \{\text{\# home simulation wins}\}/50$ and 
@@
 
-Overall, the Bayesian Network performed well, and was able to produce realistic win probabilities. Because it defined a joint distribution using a DAG and
-then used that distribution to fit the data, it was able to correctly predict accuracies using the few features provided. However, the Bayesian Network struggled to capture the intricate dependencies
-of factors other than goals that could affect the win probabilities, so the predicted probabilities are little more than an over probability calculation of goals and time reamining given historical data.
-We see this in the probability graph below, where the changes in probability correspond to the goals scored in the game.
-A sample win probability graph is seen below in Figure~\ref{fig:bayesian_graph}.
+Overall, the Bayesian network performed well, and was able to produce realistic win probabilities. Because it defined a joint distribution using a DAG and
+then used that distribution to fit the data, it was able to correctly predict accuracies using the few features provided. However, the Bayesian network struggled to capture the intricate dependencies
+of factors other than goals that could affect the win probabilities, so the predicted probabilities are little more than an over all probability calculation of goals and time reamining given historical data.
+We see this in the probability graph in Figure~\ref{fig:bayesian_graph}, where the changes in probability correspond to the goals scored in the game.
 

### 2024-04-12 Latex/draft.tex
typos and graph

@@
 did not remove any strange events from out dataset.
+
+The models were each trained on prior seasons, and tested on the most recent 2022-2023 season.
+This split was chosen because in practice our model would have access to historical data but not have access to the current season's data as it happens.
 
@@
 
-As mentioned in the methods section, we trained an XGBoost model for each team in the NHL. One main benefit of this
-approach is the ability to leverage the {\tt .predict\_proba()} method of XGBoost.
+For XGBoost, we leveraged the {\tt .predict\_proba()} method to predict win probabilities after the model was fit on the data.
 Using this method, we generated 
@@
 (see Figure~\ref{fig:xgboost_res}). 
-This approach provided insights into momentum shifts (when one team 
+This approach allowed for more insights into momentum shifts (when one team 
 has an advantage due to them having more players on the ice due to 
-a penalty by the opposing team), key moments (such as goals and 
-penalties, and critical plays influencing game outcomes).
+a penalty by the opposing team), and key moments (such as goals and 
+penalties, and critical plays influencing game outcomes) than the Bayesian network, resulting in more variance in the win probability graph.
 
@@
 To evaluate our models, we randomly selected 100 games outside of the training data. 
-To ensure that the test set represented the actual data, we selected the test set with 50 home wins and 50 away wins, and we also made sure there was an even representation of each of the 32 NHL teams.
+To ensure that the test set represented the actual data, we selected the test set with 50 home wins and 50 away wins, 
+and we also made sure there was an even representation of each of the 32 NHL teams. 
 We then used each model to predict a winner at various points throughout each game. 
@@
 
-In Figure~\ref{tab:game_accuracies} we note that each model increased in accuracy as the game progressed.
-We also notice that the MCMC simulation method appears to have the highest accuracy after the start of the game while XGBoost appears to be the least accurate.
-We infer that the XGBoost models were overfit, since they appeared to perform very well on games within the training set.
-Also, during experimentation with the models, we noticed that the MCMC simulator tended to produce higher probabilities for the away team winning, and the XGBoost model tended to give higher probabilities to the home team winning.
+In Table~\ref{tab:game_accuracies} we note that each model increased in accuracy as the game progressed. 
+The Bayesian network and MCMC model were the most accurate, while the XGBoost model had lower accuracy.
+We infer that the XGBoost models were overfit, as on the training data it would regularly achieve higher than 90\% accuracy at most points during the game. 
+Also, during experimentation with the models, we noticed that the MCMC simulator tended to produce higher probabilities for the away team winning, 
+and the XGBoost model tended to give higher probabilities to the home team winning. 
 The Bayesian Network was unbiased towards the home or away team. 
@@
 We used a Bayesian Network to model the conditional dependencies of key features, XGBoost to model the play-by-play dynamics of each team, and an MCMC simulator to simulate the flow of events in a hockey game.
-We found that the MCMC simulator was the most accurate, with the Bayesian Network coming in second, and XGBoost coming in last.
+We found that the Bayesian network and MCMC simulator were more accurate than XGBoost, and Bayesian network also was significantly faster than the MCMC simulator.
+Therefore we can conclude that the Bayesian network is the best performer for this project, implying that the conditional dependencies and potential causal relationships between features are important for predicting the outcome of hockey games.
 The win percentages predicted are more for entertainment and educational purposes, and should not be used for gambling or performance purposes, as they do not have a incredibly high degree of accuracy.
@@
 Our project was limited by the data
-we had access to and computational resources, but nonetheless we were able to produce a model that can predict the outcome of NHL games with a reasonable degree of accuracy.
+we had access to and computational resources, but nonetheless we were able to produce a model that can predict the outcome of NHL games with a reasonable degree of accuracy depending on the time of game remaining.
 

### 2024-04-12 Latex/draft.tex
Pushing The Good Changes(TM)

@@
 
-The models were each trained on prior seasons, and tested on the most recent 2022-2023 season.
+The models were each trained on prior seasons, and tested on the most recent 2023-2024 season.
 This split was chosen because in practice our model would have access to historical data but not have access to the current season's data as it happens.
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{images/good_bayesian_example.png}
+    \includegraphics[width=\textwidth]{images/good_bayesian_example.png}
     \caption{Win probability graph using the Bayesian Network. As seen above, the win probability stayed rather stagnant until a goal was scored, where the probability shifted.}
@@
     \centering
-    \subfigure[]{\includegraphics[width=0.6\textwidth]{images/good_xgb_example.png}} 
-    \subfigure[]{\includegraphics[width=0.6\textwidth]{images/good_xgb_example(2).png}} 
-    \caption{For both of these plots, the home team is above the the 50\% line, with the
-    away team below. The title has the final score of the game. \\
-        \textbf{(a)} The live win probabilities for the Nashville Predators at the Vegas Golden Knights. 
-    Notice the probability shifts after one team scores a goal.  \\
-        \textbf{(b)} The live win probabilities for the Calgary Flames at the Winnipeg Jets.
-        Notice the stark changes in probabilities after each goal. Also notice the
-        shift in probabilities even when no goals are scored.}
+    \includegraphics[width=\textwidth]{images/good_xgb_example.png}
+    \caption{We can see a high variance of win probability during this game. It is interesting to
+    see how a goal being scored impacts the win probability. It is also interesting to see how
+    the probability shifts without a goal being scored, implying the other events in the game
+    are important to the win probability.}
     \label{fig:xgboost_res}
@@
 We trained the MCMC model on the flow of events in the game, so it appears that the model picked up on a more predictive flow in how away teams play. 
-Though the MCMC model was generally the most accurate, it takes about 2 minutes to simulate 50 games and get a single prediction while XGBoost and the Bayesian Network can make a prediction in under two seconds.
+Though the MCMC model was generally the most accurate, it takes about 2 minutes to simulate 50 games and get a single prediction while XGBoost and the Bayesian network can make a prediction in under two seconds.
 

### 2024-04-12 Latex/draft.tex
Pushing final read through.

@@
 Our project seeks a similar outcome to the research mentioned above: predict the outcomes of NHL games. Not only this,
-but we seek to provide live, acturate, win probabilities for any given game state. Despite the simplicity of the problem statement, 
+but we seek to provide live, accurate, win probabilities for any given game state. Despite the simplicity of the problem statement, 
 the solution is not so straightforward. The NHL provides fast-paced games with many events
@@
 
-The models were each trained on prior seasons, and tested on the most recent 2023-2024 season.
+The models were each trained on the 2021-2023 seasons (totaling two seasons), and tested on the most recent 2023-2024 season.
 This split was chosen because in practice our model would have access to historical data but not have access to the current season's data as it happens.
@@
 events that occurred. For example, if the home team won a faceoff, lost the puck, and 
-then the away team shot the puck and missed, the markov chain would look like Table~\ref{tab:markov_chain_sample}.
+then the away team shot the puck and missed, the next potential event in the Markov chain would look like Table~\ref{tab:markov_chain_sample}.
 
@@
 
-To simulate a game, we performed this Monte Carlo Algorithm (see Algorithm~\ref{alg:monte_carlo_algorithm} in Appendix~\ref{mcmc_app})
-that acts as a random walk through the Markov Chain (Note that the {\tt KDE\_times()} is a KDE model fit on the amount of seconds that transpired between 
+To simulate a game, we performed a Monte Carlo algorithm (see Algorithm~\ref{alg:monte_carlo_algorithm} in Appendix~\ref{mcmc_app})
+that acts as a random walk through the Markov chain (note that the {\tt KDE\_times()} is a KDE model fit on the amount of seconds that transpired between 
 each hockey event for all NHL games over the course of 13 years).
@@
 
-Overall, the Bayesian network performed well, and was able to produce realistic win probabilities. Because it defined a joint distribution using a DAG and
+Overall, the Bayesian network performed well and was able to produce realistic win probabilities. Because it defined a joint distribution using a DAG and
 then used that distribution to fit the data, it was able to correctly predict accuracies using the few features provided. However, the Bayesian network struggled to capture the intricate dependencies
@@
 if these algorithms clustered the synthetic games and actual games in 
-different clusters. As shown below in Figure~\ref{fig:simulation_v_actual}, the simulated
+different clusters. As shown in Figure~\ref{fig:simulation_v_actual}, the simulated
 games and actual games are all clustered together. From these two experiments, we conclude
@@
 \subsection{Method Comparison}
-Each of our methods had one general purpose: predict the winner of the game regardless of the point of the game. 
+Each of our methods have one general purpose: predict the winner of the game regardless of the point of the game. 
 To evaluate our models, we randomly selected 100 games outside of the training data. 

### 2024-04-13 Latex/draft.tex
Found another typo.

@@
 \subsection{Method Comparison}
-Each of our methods have one general purpose: predict the winner of the game regardless of the point of the game. 
+Each of our methods has one general purpose: predict the winner of the game regardless of the point of the game. 
 To evaluate our models, we randomly selected 100 games outside of the training data. 
@@
 
-Additionally, there are no privacy concerns with our data as we only collected publicly availalbe NHL data.
+Additionally, there are no privacy concerns with our data as we only collected publicly available NHL data.
 

### 2024-04-14 Latex/draft.tex
syncing

--- 2024-04-12 Latex/draft.tex
@@
 
-Hockey; however, is a different story. While there are some methods used to predict the outcome of National Hockey League (NHL) games, these models
+Hockey, however, is a different story. While there are some methods used to predict the outcome of National Hockey League (NHL) games, these models
 typically belong to sport books and their nuances are not publicly disclosed. Additionally, hockey analytics is not as
-developed as it is in other sports, such as basketball or baseball~\cite{Brooks-Davis}. This lack of model transparency and public interest in hockey analytics
+developed as it is in other sports, such as basketball or baseball. This lack of model transparency and public interest in hockey analytics
 makes predicting the outcomes of NHL games a very underdeveloped and challenging problem. Previous attempts and research into predicting NHL games
@@
 
-In addition to model research, some effort has also gone into developing new features that can be used to better predict the game outcomes. 
+In addition to model research, some research has also gone into developing new features that can be used to better predict the game outcomes. 
 The two biggest engineered classes of features are the Corsi
@@
  We were unable to locate the original blog posts talking about these metrics, but a good article to learn more about the math can be
- found here \url{https://thehockeywriters.com/corsi-fenwick-stats-what-are-they/}.} metrics (both around 2007); however, we do not use these
- in our approach because future research shows they do not improve model performance~\cite{rinkstats_corsi_fenwick}.
+ found here \url{https://thehockeywriters.com/corsi-fenwick-stats-what-are-they/}.} metrics (both around 2007).
 
 Our project seeks a similar outcome to the research mentioned above: predict the outcomes of NHL games. Not only this,
-but we seek to provide live, acturate, win probabilities for any given game state. Despite the simplicity of the problem statement, 
-the solution is not so straightforward. The NHL provides fast-paced games with many events
+but we seek to provide live, acturate win probabilities for any given game state. Despite the simplicity of the problem statement, 
+as mentioned, the solution is not so straightforward. The NHL provides fast-paced games with many events
 occuring in quick succession. Our goal is to use this abundance of data and new approaches to build upon previous research.
@@
 % Your content for this section goes here
-Our data came from the hockeyR Github repository~\cite{hockeyR-data}. This repository contains an abundance of data about every NHL game
-that has occured since the 2010-11 season. This data includes information about the events that transpire in a game (hits, shots, goals, etc.),
+Our data came from the hockeyR Github repository~\cite{hockeyR-data}. With data starting in the 2010-2011 season, this dataset contains events that transpire in a game (hits, shots, goals, etc.),
 which teams are playing, who is on the ice, and the final score of the game. The data is stored in a series of {\tt .csv.gz} files, allowing for
@@
 Each game in a season is given a unique identifier ({\tt game\_id}), which is constant across all events in a game. Every event that occurs in a game
-will be stored in the {\tt event\_type} column. There are 17 unique event types, including game start, faceoff, shot, hit, and goal.
+will be stored in the {\tt event\_type} column. There are 17 unique event types, including things such as game start, faceoff, shot, hit, and goal.
 Most of these event types are not relevant to our analysis, so we remove them from the dataset. After removing the unnecessary events, we are left with
-nine: blocked shot, faceoff, giveaway, goal, hit, missed shot, penalty, shot, and takeaway. These events are attributed to the
-team and player that perform the event. We only take into consideration the team that performs the event and discard the player information to reduce noise and to avoid many one-hot encoded columns.
+nine events: blocked shot, faceoff, giveaway, goal, hit, missed shot, penalty, shot, and takeaway. These events are attributed to the
+team and player that perform the event, but we only take into consideration the team that performs the event and discard the player information to reduce 
+noise and to avoid many one-hot encoded columns.
 
@@
 
-We found that our data did not contain any missing values that were not easily explainable. For example, if a game is starting, there will be no
+We found that our data did not contain any missing values that was not easily explainable. For example, if a game is starting, there will be no
 events for penalties, which will result in a {\tt NaN} value in the penalties column. Additionally, any data that was confusing or not easily explainable
@@
 
-The models were each trained on prior seasons, and tested on the most recent 2023-2024 season.
-This split was chosen because in practice our model would have access to historical data but not have access to the current season's data as it happens.
-
 \section{Methods}
@@
 \subsection{Bayesian Network}
-We first used a Bayesian network to establish a benchmark for probability using several key features.
+We first used a Bayesian Network to establish a benchmark for probability using several key features.
 
@@
 The computational complexity of Bayesian Network inference is high, with exact inference being an NP-hard problem~\cite{pmlr-vR0-chickering95a}. 
-Using the python package {\tt pgmpy}, we originally tried to fit a network with all 26 of our features, but our computational resources failed to fit this network.
+Using the python package pgmpy, we originally tried to fit a network with all 26 of our features, but our computational resources failed to fit this network.
 Then, to get a baseline for our future predictions, we simply fitted the model with the base features of time remaining (tr), home goals (hg), away goals (ag), home shots (hs), away shots (as), home blocked shots (hbs), and away blocked shots (abs) in order to predict wins (w).
@@
 
-XGBoost was chosen because of its ability to handle large datasets and model complex relationships between features. 
-XGBoost is able to attain a high degree of accuracy with a relatively small training time or overhead and was an ideal choice for our research.
-
 \subsection{MCMC Game Simulation}
@@
 To simulate hockey games, we created a Markov Chain where the states are a tuple of three consecutive
-events that occurred. For example, if the home team won a faceoff, lost the puck, and 
+events that occurred. For example, if the home team won a faceoff, the home team lost the puck, and 
 then the away team shot the puck and missed, the markov chain would look like Table~\ref{tab:markov_chain_sample}.
@@
 
-To simulate a game, we performed this Monte Carlo Algorithm (see Algorithm~\ref{alg:monte_carlo_algorithm} in Appendix~\ref{mcmc_app})
-that acts as a random walk through the Markov Chain (Note that the {\tt KDE\_times()} is a KDE model fit on the amount of seconds that transpired between 
+To simulate a game, we performed this Monte Carlo Algorithm (see Algorithm~\ref{alg:monte_carlo_algorithm} in Appendix~\ref{mcmc_app}) that acted like a random walk through the Markov Chain (Note that the {\tt KDE\_times()} is a KDE model fit on the amount of seconds that transpired between
 each hockey event for all NHL games over the course of 13 years).
@@
 
-To predict a team's winning probability, we would simulate 50 hockey games with the initial starting states 
-$(s_0, s_1, s_2)$ set to the most recent events in the hockey game. By looking at each games
-final event counts we compute the winner. We then compute the 
+To predict a team's winning probability, we would simulate 50 hockey games with the initial starting states $(s_0, s_1, s_2)$ set to the most current events in the hockey game.
+By looking at each games
+final event counts, we could compute each game's winner. Then we computed the 
 probability as $P(\text{home winning}) = \{\text{\# home simulation wins}\}/50$ and 
@@
 
-Overall, the Bayesian network performed well, and was able to produce realistic win probabilities. Because it defined a joint distribution using a DAG and
-then used that distribution to fit the data, it was able to correctly predict accuracies using the few features provided. However, the Bayesian network struggled to capture the intricate dependencies
-of factors other than goals that could affect the win probabilities, so the predicted probabilities are little more than an over all probability calculation of goals and time reamining given historical data.
-We see this in the probability graph in Figure~\ref{fig:bayesian_graph}, where the changes in probability correspond to the goals scored in the game.
+Overall, the Bayesian Network performed well, and was able to produce realistic win probabilities. Because it defined a joint distribution using a DAG and
+then used that distribution to fit the data, it was able to correctly predict accuracies using the few features provided. However, the Bayesian Network struggled to capture the intricate dependencies
+of factors other than goals that could affect the win probabilities, so the predicted probabilities are little more than an over probability calculation of goals and time reamining given historical data.
+We see this in the probability graph below, where the changes in probability correspond to the goals scored in the game.
+A sample win probability graph is seen below in Figure~\ref{fig:bayesian_graph}.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{images/good_bayesian_example.png}
+    \includegraphics[width=0.9\textwidth]{images/good_bayesian_example.png}
     \caption{Win probability graph using the Bayesian Network. As seen above, the win probability stayed rather stagnant until a goal was scored, where the probability shifted.}
@@
 
-For XGBoost, we leveraged the {\tt .predict\_proba()} method to predict win probabilities after the model was fit on the data.
+As mentioned in the methods section, we trained an XGBoost model for each team in the NHL. One main benefit of this
+approach is the ability to leverage the {\tt .predict\_proba()} method of XGBoost.
 Using this method, we generated 
@@
 (see Figure~\ref{fig:xgboost_res}). 
-This approach allowed for more insights into momentum shifts (when one team 
+This approach provided insights into momentum shifts (when one team 
 has an advantage due to them having more players on the ice due to 
-a penalty by the opposing team), and key moments (such as goals and 
-penalties, and critical plays influencing game outcomes) than the Bayesian network, resulting in more variance in the win probability graph.
+a penalty by the opposing team), key moments (such as goals and 
+penalties, and critical plays influencing game outcomes).
 
@@
     \centering
-    \includegraphics[width=\textwidth]{images/good_xgb_example.png}
-    \caption{We can see a high variance of win probability during this game. It is interesting to
-    see how a goal being scored impacts the win probability. It is also interesting to see how
-    the probability shifts without a goal being scored, implying the other events in the game
-    are important to the win probability.}
+    \subfigure[]{\includegraphics[width=0.6\textwidth]{images/good_xgb_example.png}} 
+    \subfigure[]{\includegraphics[width=0.6\textwidth]{images/good_xgb_example(2).png}} 
+    \caption{For both of these plots, the home team is above the the 50\% line, with the
+    away team below. The title has the final score of the game. \\
+        \textbf{(a)} The live win probabilities for the Nashville Predators at the Vegas Golden Knights. 
+    Notice the probability shifts after one team scores a goal.  \\
+        \textbf{(b)} The live win probabilities for the Calgary Flames at the Winnipeg Jets.
+        Notice the stark changes in probabilities after each goal. Also notice the
+        shift in probabilities even when no goals are scored.}
     \label{fig:xgboost_res}
@@
 To evaluate our models, we randomly selected 100 games outside of the training data.
-To ensure that the test set represented the actual data, we selected the test set with 50 home wins and 50 away wins, 
-and we also made sure there was an even representation of each of the 32 NHL teams. 
+To ensure that the test set represented the actual data, we selected the test set with 50 home wins and 50 away wins, and we also made sure there was an even representation of each of the 32 NHL teams.
 We then used each model to predict a winner at various points throughout each game.
@@
 
-In Table~\ref{tab:game_accuracies} we note that each model increased in accuracy as the game progressed. 
-The Bayesian network and MCMC model were the most accurate, while the XGBoost model had lower accuracy.
-We infer that the XGBoost models were overfit, as on the training data it would regularly achieve higher than 90\% accuracy at most points during the game. 
-Also, during experimentation with the models, we noticed that the MCMC simulator tended to produce higher probabilities for the away team winning, 
-and the XGBoost model tended to give higher probabilities to the home team winning. 
+In Figure~\ref{tab:game_accuracies} we note that each model increased in accuracy as the game progressed.
+We also notice that the MCMC simulation method appears to have the highest accuracy after the start of the game while XGBoost appears to be the least accurate.
+We infer that the XGBoost models were overfit, since they appeared to perform very well on games within the training set.
+Also, during experimentation with the models, we noticed that the MCMC simulator tended to produce higher probabilities for the away team winning, and the XGBoost model tended to give higher probabilities to the home team winning.
 The Bayesian Network was unbiased towards the home or away team.
 We trained the MCMC model on the flow of events in the game, so it appears that the model picked up on a more predictive flow in how away teams play.
-Though the MCMC model was generally the most accurate, it takes about 2 minutes to simulate 50 games and get a single prediction while XGBoost and the Bayesian network can make a prediction in under two seconds.
+Though the MCMC model was generally the most accurate, it takes about 2 minutes to simulate 50 games and get a single prediction while XGBoost and the Bayesian Network can make a prediction in under two seconds.
 
@@
 We used a Bayesian Network to model the conditional dependencies of key features, XGBoost to model the play-by-play dynamics of each team, and an MCMC simulator to simulate the flow of events in a hockey game.
-We found that the Bayesian network and MCMC simulator were more accurate than XGBoost, and Bayesian network also was significantly faster than the MCMC simulator.
-Therefore we can conclude that the Bayesian network is the best performer for this project, implying that the conditional dependencies and potential causal relationships between features are important for predicting the outcome of hockey games.
+We found that the MCMC simulator was the most accurate, with the Bayesian Network coming in second, and XGBoost coming in last.
 The win percentages predicted are more for entertainment and educational purposes, and should not be used for gambling or performance purposes, as they do not have a incredibly high degree of accuracy.
@@
 Our project was limited by the data
-we had access to and computational resources, but nonetheless we were able to produce a model that can predict the outcome of NHL games with a reasonable degree of accuracy depending on the time of game remaining.
+we had access to and computational resources, but nonetheless we were able to produce a model that can predict the outcome of NHL games with a reasonable degree of accuracy.
 

### 2024-04-14 Latex/draft.tex
syncing

--- 2024-04-13 Latex/draft.tex
@@
 We use each of these methods to generate win probabilities for each time step throughout various NHL games, thus emulating a live win probability.
-Finally, we analyze the accuracy of these three methods and considered ethical implications of our results.
+Finally, we analyze the accuracy of these three methods and consider ethical implications of our results.
 \end{abstract}
@@
 developed as it is in other sports, such as basketball or baseball~\cite{Brooks-Davis}. This lack of model transparency and public interest in hockey analytics
-makes predicting the outcomes of NHL games a very underdeveloped and challenging problem. Previous attempts and research into predicting NHL games
+makes predicting the outcomes of NHL games a very underdeveloped and challenging problem. Previous attempts and research in predicting NHL games
 has relied on methods such as decision trees and artificial neural networks~\cite{pishcedda} (from 2014), naïve bayes and support vector machines~\cite{weissbock2013use} (from 2013),
@@
 % Your content for this section goes here
-Our data came from the hockeyR Github repository~\cite{hockeyR-data}. This repository contains an abundance of data about every NHL game
+Our data comes from the hockeyR Github repository~\cite{hockeyR-data}. This repository contains an abundance of data about every NHL game
 that has occured since the 2010-11 season. This data includes information about the events that transpire in a game (hits, shots, goals, etc.),
@@
 
-Each game in a season is given a unique identifier ({\tt game\_id}), which is constant across all events in a game. Every event that occurs in a game
-will be stored in the {\tt event\_type} column. There are 17 unique event types, including game start, faceoff, shot, hit, and goal.
-Most of these event types are not relevant to our analysis, so we remove them from the dataset. After removing the unnecessary events, we are left with
-nine: blocked shot, faceoff, giveaway, goal, hit, missed shot, penalty, shot, and takeaway. These events are attributed to the
-team and player that perform the event. We only take into consideration the team that performs the event and discard the player information to reduce noise and to avoid many one-hot encoded columns.
+Each game in a season has a unique identifier ({\tt game\_id}), which is constant across all events in a game.
+Every event that occurs in a game is stored in the {\tt event\_type} column.
+Most of the 17 unique event types are not relevant to our analysis, so we only include the prominent ones: blocked shot, faceoff, giveaway, goal, hit, missed shot, penalty, shot, and takeaway.
+We also leave out the player information for each event to reduce noise and to avoid many one-hot encoded columns.
 
-The data also contains information about when the event occured. This appears in a variaty of formats, but we only
+The data also contains information about when the event occurred. This appears in a variety of formats, but we only
 use the {\tt game\_time\_remaining} column. {\tt game\_time\_remaining} starts
-at 3600 (60 minutes) and counts down to 0. If the game goes into extra time, i.e., it is tied after 60 minutes, {\tt game\_time\_remaining} will
+at 3600 (60 minutes) and counts down to 0. If the game goes into extra time (i.e. it is tied after 60 minutes), {\tt game\_time\_remaining} will
 be a negative value.
@@
 The models were each trained on the 2021-2023 seasons (totaling two seasons), and tested on the most recent 2023-2024 season.
-This split was chosen because in practice our model would have access to historical data but not have access to the current season's data as it happens.
+This split ensures that the models use the most relevant data for training while still leaving the most recent season as testing to simulate live prediction probabilities.
 
@@
 To simulate a game, we performed a Monte Carlo algorithm (see Algorithm~\ref{alg:monte_carlo_algorithm} in Appendix~\ref{mcmc_app})
-that acts as a random walk through the Markov chain (note that the {\tt KDE\_times()} is a KDE model fit on the amount of seconds that transpired between 
-each hockey event for all NHL games over the course of 13 years).
+that acts as a random walk through the Markov chain. Note that the {\tt KDE\_times()} in the algorithm is a KDE model fit on the amount of seconds that transpired between 
+each hockey event for all NHL games over the course of 13 years.
 % \begin{algorithm}
@@
 
-To predict a team's winning probability, we would simulate 50 hockey games with the initial starting states 
-$(s_0, s_1, s_2)$ set to the most recent events in the hockey game. By looking at each games
-final event counts we compute the winner. We then compute the 
+To predict a team's winning probability, we simulate 50 hockey games with the initial starting states 
+$(s_0, s_1, s_2)$ set to the most recent events in the hockey game. By looking at each game's final event counts, we compute the winner. We then compute the 
 probability as $P(\text{home winning}) = \{\text{\# home simulation wins}\}/50$ and 
@@
 Overall, the Bayesian network performed well and was able to produce realistic win probabilities. Because it defined a joint distribution using a DAG and
-then used that distribution to fit the data, it was able to correctly predict accuracies using the few features provided. However, the Bayesian network struggled to capture the intricate dependencies
-of factors other than goals that could affect the win probabilities, so the predicted probabilities are little more than an over all probability calculation of goals and time reamining given historical data.
+then used that distribution to fit the data, it was able to correctly predict accuracies using the few features provided.
+However, the Bayesian network struggled to capture the intricate dependencies of factors other than goals that could affect the win probabilities, so the predicted probabilities are little more than an over all probability calculation of goals and time remaining given historical data.
 We see this in the probability graph in Figure~\ref{fig:bayesian_graph}, where the changes in probability correspond to the goals scored in the game.
